### 2015-05-29 LSP/lsp-seriesinfo/cmle-info.tex
initial commit paper

%% -*- coding:utf-8 -*-
{\large Computational Models of Language Evolution}

\bigskip

Editors: Luc Steels, Remi van Trijp

\bigskip

In this series:

\begin{enumerate}
\item Steels, Luc. The Talking Heads experiment: origins of words and meanings.
\item Vogt, Paul. How mobile robots can self-organize a vocabulary.
\item Bleys, Joris. Language strategies for the domain of colour.
\item van Trijp, Remi. The evolution of case grammar.
\item Spranger, Michael. The evolution of grounded spatial language.
\end{enumerate}

% 
% \vfill
% 
% \gdef\lsISSN{2363-5568}%
% \hfill ISSN: \lsISSN
%