### 2015-05-29 chapters/pitfalls.tex

\chapter{Unicode Pitfalls}
\label{unicode-pitfalls}

\section{Wrong it is not}
\label{wrong-it-is-not}

In this section, we describe some of the most common pitfalls that we have encountered when using the Unicode Standard in our own work, or in discussion with other linguists. This section is not meant as a criticism of the decisions made by the Unicode Consortium; on the contrary, we aim to highlight where the technological aspects of the Unicode Standard diverge from many users' intuitions. What have sometimes been referred to as problems or inconsistencies in the Unicode Standard are mostly due to legacy compatibility issues, which can lead to unexpected behavior by linguists using the standard. However, there are also some cases in which the Unicode Standard has made decisions that theoretically could have been taken differently, but for some reason or another (mostly very good reasons) were accepted as they are now. We call behavior that executes without error but does something different than the user expected---often unknowingly---a ``pitfall''.

In this context, it is important to realize that the Unicode Standard was not developed to solve linguistic problems per se, but to offer a consistent computational environment for written language. In those cases in which the Unicode Standard behaves differently as expected, we think it is important not to dismiss Unicode as `wrong' or `deficient', because our experience is that in almost all cases the behaviour of the Unicode Standard has been particularly well thought through. The Unicode Consortium considers important practical use-cases that from a linguistic point of view are normally not considered. Our general guideline for dealing with the Unicode Standard is to accept it as it is, and not to battle windmills. Alternatively, of course, it is possible to actively engage in the development of the standard itself, an effort that is highly appreciated by the Unicode Consortium.

\section{Pitfall: Blocks}
\label{pitfall-blocks}

The Unicode code space is subdivided into blocks of contiguous code points. For example, the block called ``Cyrillic'' runs from U+0400 till U+04FF. These blocks arose as an attempt at ordering the enormous amount of characters in Unicode, but the ideas of blocks very quickly ran into problems. First, the size of a block is fixed, so when a block is full, a new block will have to be instantiated somewhere further in the code space. For example, this led to the blocks ``Cyrillic Supplement'', ``Cyrillic Extended-A'' (which are also already full) and ``Cyrillic Extended-B''. Second, when a specific character already exists, then it is not duplicated in another block, although the name of the block might indicate that a specific symbol should be available there. In general, names of blocks are just an approximate indication of the kind of characters that will be in the block.

The problem with blocks arises because finding the right character among the thousands of Unicode characters is not easy. Many software applications present blocks as a primary search mechanism, because the block names suggest where to look for a particular character. However, when a user searches for an IPA character in the block ``IPA Extensions'' then many IPA characters will not be found there. For example, the velar nasal <ŋ> is not part of the block ``IPA Extensions'' because it was already included as LATIN SMALL LETTER ENG at U+014B in the block ``Latin Extensions-A''.

In general, finding a specific character in the Unicode Standard is often not trivial. The names of the blocks can help, but it is not (and never was supposed to be) a foolproof structure. It is not the goal nor aim of the Unicode Consortium to provide a user interface to the Unicode Standard. If one often encounters the problem of needing to find a suitable character, there are various other useful services for end users available.\footnote{The Unicode website offers a basic interface to the code charts at http://www.unicode.org/charts/index.html. As a more flexible interface, we particularly like PopChar from Macility, available for both Macintosh and Windows. However, there are various free websites that offer search interfaces to the Unicode code tables, like http://unicode-search.net or http://unicode-search.net. A further useful approach for searching characters using shape matching is http://shapecatcher.com.}

\section{Pitfall: Missing glyphs}
\label{pitfall-missing-glyphs}

The Unicode Standard is often praised (and deservedly so) for solving many of the perennial problems with the interchange and display of the world's writing systems. However, a common complaint from users is that, while the praise may be true, they mostly just see some boxes on their screen instead of those promised symbols. The problem of course is that users' computers do not have any glyphs installed matching the Unicode code points in the file they are trying to inspect. It is important to realize that internally in the computer everything still works as expected: any handling of Unicode code points works independently of how they are displayed on the screen. So, although a user might only see boxes being displayed, this user should be assured that everything is still in order.

The central problem behind the missing glyphs is that designing actual glyphs includes a lot of different considerations and it is a time-consuming process. Many traditional expectations of how specific characters should look like have to be taken into account when designing glyphs. Those expectations are often not well documented, and it is mostly up to the knowledge and experience of the font designer to try and conform to them as good as possible. Therefore, most designers produce fonts only including glyphs for certains parts of the Unicode Standard, namely for those characters they feel comfortable with. At the same time, the number of characters defined by the Unicode Standard is growing with each new version, so it is neigh impossible for any designer to produce glyphs for all characters. The result of this is that, almost necessarily, each font only includes glyphs for a subset of the characters in the Unicode Standard.

The simple solution to missing glyphs is thus to install additional fonts providing additional glyphs. For the more exotic characters there is often not much choice. There are a few particularly large fonts that might be considered. First, there is the \textsc{Everson Mono} font made by Michael Everson, which currently includes 9,756 different glyphs (not including Chinese) updated up to Unicode 7.0.\footnote{Everson Mono is available as shareware at http://www.evertype.com/emono/.} Already a bit older is the \textsc{Titus Cyberbit Basic} font made by Jost Gippert and Carl-Martin Bunz, which includes 10,044 different glyphs (not including Chinese), but not including newer characters added after Unicode 4.0.\footnote{Titus Cyberbit Basic is available at http://titus.fkidg1.uni-frankfurt.de/unicode/tituut.asp.}

Further, we suggest to always install at least one so-called \textsc{fall-back font,} which provides glyphs that at least show the user some information about the underlyingly encoded character. Apple Macintoshes have such a font (which is invisible to the user), which is designed by Michael Everson and made available for other systems through the Unicode Consortium.\footnote{The Apple/Everson fallback font is available at http://www.unicode.org/policies/lastresortfont\_eula.html.} Further, the \textsc{GNU Unifont} is a clever way to produce bitmaps approximating the intended glyph of each available character, updated to Unicode 7.0.\footnote{The GNU Unifont is available at http://unifoundry.com/unifont.html.} Finally, the Summer Institute of Linguistics provides a \textsc{SIL Unicode BMP Fallback Font}, currently available up to Unicode version 6.1. This font does not even attempt to show a real glyph, but only shows the hexadecimal code inside a box for each character, so a user can at least see the Unicode codepoint of the character to be displayed.\footnote{The SIL Unicode BMP Fallback Font is available at http://scripts.sil.org/UnicodeBMPFallbackFont.}

\section{Pitfall: Faulty rendering}
\label{pitfall-faulty-rendering}

A similar complaint to missing glyphs, discussed previously, is that while there might be a glyph being displayed, it does not `look right'. There are two reasons for unexpected visual display, namely automatic font substitution and faulty rendering. Like missing glyphs, any such problems are independent from the Unicode Standard. The Unicode Standard only includes very general information about characters and leaves the specific visual display to others to decide on. Any faulty display is thus not to be blamed on the Unicode Consortium, but on a complex interplay of different mechanisms happening in a computer to turn Unicode codepoints into visual symbols. We will only sketch a few aspects of this complex interplay here.

Most modern software applications (like Microsoft Word) offer some approach to \textsc{automatic font substitution}. This means that when a text is written in a specific font (e.g.~Times New Roman) and a inserted Unicode character does not have a glyph within this font, then the software application will automatically search for another font to display the glyph. The result will be that some specific glyph will look slightly different from the others. This mechanism works differently depending on the software application, and mostly only limited user influence is expected and little feedback is given, which might be rather frustrating to font-aware users.\footnote{For example, Apple Pages does not give any feedback that a font is being replaced, and the user does not seem to have any influence on the choice of replacement (except by manually marking all occurrences). In contrast, Microsoft Word does indicate the font replacement by showing the name of the font replacement, but it simply changes the font, so any text written after the replacement is written in a different font as before. Both behaviors leave much to be desired.}

The other problem with visual display is related to the so-called `font rendering'. \textsc{Font rendering }refers to the process of the actual positioning of Unicode characters on a page of written text. This positioning is actually a highly complex problem, and many things can go wrong in the process. Well-known rendering problems, like proportional glyph size or ligatures are reasonably well understood. However, even within the Latin script, the positioning of diacritics relative to a base character is still a widespread problem. Especially when more than one diacritic is supposed to be placed above (or below) each other, this often leads to unexpected effects in many modern software applications. The problems arising in Arabic and in many southeast Asian scripts (like Devanagari or Burmese) are even much more complex. There are basically three different approaches to font rendering. The most widespread is Adobe's and Microsoft's \emph{OpenType }system. This approach makes it relatively easy for font developers, as the font itself does not include all details about the precise placement of individual characters. For those details, additional script-descriptions are necessary. All of those systems can lead to unexpected behaviour.\footnote{For more details about OpenType, see http://www.adobe.com/products/type/opentype.html and http://www.microsoft.com/typography/otspec/. Additional systems for complex text layout are, among others, Microsoft's DirectWrite https://msdn.microsoft.com/library/dd368038.aspx and the open-source project HarfBuzz http://www.freedesktop.org/wiki/Software/HarfBuzz/.} Alternative systems are Apple's \emph{Apple Advanced Typography} (AAT) and the open-source \emph{Graphite} system from the Summer Institute of Linguistics (SIL).\footnote{More information about AAT can be found at https://developer.apple.com/fonts/. SIL's Graphite is described in detail at http://scripts.sil.org/cms/scripts/page.php?site\_id=projects\&item\_id=graphite\_home.} In both of these systems, a larger burden is placed on the description inside the font.

\section{Pitfall: Homoglyphs}
\label{pitfall-homoglyphs}

Homoglyphs are visually indistinguishable glyphs (or highly similar glyphs) that have different code points in the Unicode Standard and thus different character semantics. As a principle, the Unicode Standard does not specify how a character appears visually on the page or the screen. So in most cases, a different appearance is caused by the specific design of a font, or by user-settings like size or boldface. Taking an example already discussed in the previous section, the following symbols <reinsert Gs here> are different glyphs of the same character, i.e.~they may be rendered differently depending on the typography being used, but they all share the same code point (viz. LATIN SMALL LETTER G at U+0067). In contrast, the symbols <> are all different code points, although they look highly similar -- in some cases even sharing exactly the same glyph in some fonts. All these different A-like characters include the following code points in the Unicode Standard:
\begin{itemize}
	\item <A> LATIN CAPITAL LETTER A, at U+0041 
	\item <> CYRILLIC CAPITAL LETTER A, at U+0410 
	\item <> GREEK CAPITAL LETTER ALPHA, at U+0391 
	\item <> CHEROKEE LETTER GO, at U+13AA 
	\item <> CANADIAN SYLLABICS CARRIER GHO, at U+15C5 
	\item <> LATIN LETTER SMALL CAPITAL A, at U+1D00 
	\item <> LISU LETTER A, at U+A4EE 
	\item <> FULLWIDTH LATIN CAPITAL LETTER A, at U+FF21 
	\item <> CARIAN LETTER A, at U+102A0 
	\item <> OLD ITALIC LETTER A, at U+10300 
	\item <> MATHEMATICAL SANS-SERIF CAPITAL A, U+1D5A0 
	\item <> MATHEMATICAL MONOSPACE CAPITAL A, at U+1D670 
\end{itemize}

The existence of such homoglyphs is partly due to legacy compatibility, but for the most part these characters are simply different characters that happen to look similar.\footnote{A particularly nice interface to look for homoglyphs is http://shapecatcher.com, based on the principle of recognizing shapes (Belongie et al. 2002).} Yet, they are supposed to behave different from the perspective of a font designer. For example, when designing a Cyrillic font, the will have different aesthetics and different traditional expectation compared to a Latin .

Such homoglyphs are a widespread problem for consistent encoding. Although for most users it looks like the words and are identical, in actual fact they do not even share a single code point.\footnote{The first words consists completely of Latin characters, namely U+0076, U+006F, U+0063, U+0065, U+0073, while the second is a mix of Cyrillic and Greek characters, namely U+03BD, U+03BF, U+0041, U+0435, U+0455.} For computers these two words are completely different entities. Commonly, when users with Cyrillic or Greek keyboards have to type some Latin-based orthography, they mix similar looking Cyrillic or Greek characters into their text, because those characters are so much easier to type. Similarly, when users want to enter an unusual symbol, they normally search by visual impression in their favourite software application, and just pick something that looks reasonably alike to what they expect the glyph to look like.

\section{Pitfall: Canonical equivalence}
\label{pitfall-canonical-equivalence}

For some characters, there is more than one possible encoding in the Unicode Standard. This is a possible pitfall, as this would mean that for the computer there exist multiple different entities that for a user are the same. This would, for example, lead to problems with searching, as the computer would search for specific encodings, and not find all expected characters. As a solution, the Unicode Standard includes a notion of \textsc{canonical equivalence}. Different encodings are explicitly declared as equivalent in the Unicode Standard code tables. Further, to harmonize all encodings in a specific piece of text, the Unicode Standard proposes a mechanism of \textsc{normalization}.

Consider for example the characters and following Unicode code points:
\begin{enumerate}
	\def\labelenumi{\arabic{enumi}.} 
	\item LATIN CAPITAL LETTER A WITH RING ABOVE (U+00C5) 
	\item ANGSTROM SIGN (U+212B) 
	\item LATIN CAPITAL LETTER A (U+0041) + COMBINING RING ABOVE (U+030A) 
\end{enumerate}

The character, represented here by glyph , is encoded in the Unicode Standard in the first two examples by a single-character sequence; each is assigned a different code point. In the third example, the glyph is encoded in a multiple-character sequence that is composed of two character code points. All three sequences are \textsc{canonically equivalent}, i.e.~they are strings that represent the same abstract character and because they are not distinguishable by the user, the Unicode Standard requires them to be treated the same in regards to their behavior and appearance. Nevertheless, they are encoded differently. For example, if one were to search an electronic text (with software that does not apply Unicode Standard normalization) for ANGSTROM SIGN (U+212B), instances of LATIN CAPITAL LETTER A WITH RING ABOVE (U+00C5) would not be found.

In other words, there are equivalent sequences of Unicode characters that should be normalized, i.e.~transformed into a unique Unicode-sanctioned representation of a character sequence called a \textsc{normalization form}. Unicode provides a \emph{Unicode Normalization Algorithm}, which essentially puts combining marks into a specific logical order and it defines decomposition and composition transformation rules to convert each string into one of four normalization forms. We will discuss here the two most relevant normalization forms: NFC and NFD.

The first of the three characters above is considered the \textsc{Normalization Form C (NFC)}, where ``C'' stands for composition. When the process of NFC normalization is applied to the character sequences in 2 and 3, both sequences are normalized into the `pre-composed' character sequence in 1. Thus all three canonical character sequences are standardized into one composition form in NFC. The other central Unicode normalization form is the \textsc{Normalization Form D (NFD)}, where ``D'' stands for decomposition. When NFD is applied to the three examples above, all three, including importantly the single-character sequences in 1 and 2, are normalized into the `decomposed' multiple-sequence of characters in 3. Again, all three are then logically equivalent and therefore comparable and syntactically interoperable.

As illustrated, some characters in the Unicode Standard have alternative representations (in fact, many do), but the Unicode Normalization Algorithm can be used to transform certain sequences of characters into canonical (or compatible) forms to test for equivalency. To determine equivalence, each character in the Unicode Standard is associated with a combining class, which is formally defined as a character property called ``Canonical Combining Class'' which is specified in the Unicode Character Database. The combining class assigned to each code point is a numeric value between 0 and 254 and is used by the Canonical Ordering Algorithm to determine which sequences of combining marks are canonically equivalent. Normalization forms, as very briefly described above, can be used to ensure character equivalence by ordering character sequences so that they can be faithfully compared.

It is very important to note that any software applications that is Unicode Standard compliant is free to change the character stream from one representation to another. This means that a software application may compose, decompose or reorder characters as its developers desire; as long as the resultant strings are canonically equivalent to the original. This might lead to unexpected behaviour for users. Various players, like the Unicode Consortium, the W3C, or the TEI recommend NFC in most user-directed situations, and some software applications that we tested indeed seem to automatically convert strings into NFC.\footnote{See the summary of various recommendation here: http://www.win.tue.nl/\textasciitilde{}aeb/linux/uc/nfc\_vs\_nfd.html. } This means in practice that if a user, for example, enters <a> and <`>, i.e.~LATIN SMALL LETTER A at U+0061 and COMBINING GRAVE ACCENT at U+0300, this might be converted into <à>, i.e.~LATIN SMALL LETTER A WITH GRAVE at U+00E0.\footnote{The behaviour of software applications can be quite erratic in this respect. For example, Apple's TextEdit does not do any conversion on text entry. However, when you copy and paste some text inside the same document in ``rich text mode'' (i.e.~RTF-format), it will be transformed into NFC on paste. Saving a document does not do any conversion to the glyphs on screen, but it will save the characters in NFC.}

\section{Pitfall: Characters are not glyphs}
\label{pitfall-characters-are-not-glyphs}

A central principle of Unicode is the distinction between character and glyph. A character is the abstract notion of a symbol in a writing system, while a glyph is the concrete drawing of such a symbol. In practice, there is a complex interaction between characters and glyphs. A single Unicode character may of course be rendered as a single glyph. However, a character may also be a `piece' of a glyph, or vice-versa. Actually, all possible relations between glyphs and characters are attested.

First, \textsc{a single character may have different contextually determined glyphs.} For example, characters in writing systems like Hebrew and Arabic have different glyphs depending on where they appear in a word. Some letters in Hebrew change their form at the end of the word, and in Arabic, primary letters have four contextually-sensitive variants (isolated, word initial, medial and final). Second, \textsc{a single character may be rendered as a sequence of multiple glyphs.} For example, in Tamil one Unicode character may result in a combination of a consonant and vowel, which are rendered as two adjacent glyphs by fonts that supports Tamil. Third, \textsc{a single glyph may be a combination of multiple characters. }For example, the ligature <ﬁ>, a single glyph, is the result of two characters, <f> and <i>, that have undergone glyph substitution by font rendering. Like contextually-determined glyphs, ligatures are (intended) artifacts of text processing instructions. Finally, \textsc{a single glyph may be a part of a character} (e.g.~diacritics).

Further, the rendering of a glyph is also always dependent on the font being used. For example, the Unicode character LATIN SMALL LETTER G appears as and in the Computer Modern and Courier fonts because their typefaces are designed differently. Furthermore, font face may change the visual appearance of a character, for example Times New Roman two-story <a> changes to a single-story a in italics <\emph{a}>. This becomes a real problem for some phonetic typesetting (see Section 5, Pitfall 2).

In sum, character-to-glyph mappings are complex technological issues that the Unicode Consortium has had to address in the development of the Unicode Standard, but for the lay user they can be utterly confusing because visual rendering does not indicate logical encoding, or, in other words, what you see isn't (necessarily) what you get.

\section{Pitfall: File formats}
\label{pitfall-file-formats}

Unicode is a character encoding standard, but these characters of course actually appear inside some kind of file. The most basic Unicode-based file format is pure line-based text, i.e.~strings of Unicode-encoded characters separated by line breaks (note that these line breaks are what for most people intuitively corresponds to paragraph breaks). Unfortunately, even within this apparently basic setting there exist a multitude of variants. In general, these different possibilities are well-understood in the software industry, and nowadays they normally do not lead to any problems for the end user. However, there are some situations in which a user is suddenly confronted with cryptic questions in the user interface involving abbreviations like LF, CR, BE, LE or BOM. Most prominently this occurs with exporting or importing data in several software applications from Microsoft. Basically, there are two different issues involved. First, the encoding of** line breaks** and, second, the encoding of the Unicode characters into \textsc{code units} and the related issue of \textsc{endianness}.

The issue with \textsc{line breaks} originated with the instructions necessary to direct a printing head to a new line. This involves two movements, known as \textsc{carriage return (CR)} (returning the printing head to the start of the line on the page) and \textsc{line feed (LF)} (moving the printing head to the next line on the page). Physically, these are two different events, but conceptually together they form one action. In the history of computing, various encodings of line breaks have been used (e.g.~CR+LF, LF+CR, only LF, or only CR). Currently, all Unix and Unix-derived systems use only LF as code for a line break, while software from Microsoft still uses a combination of CR and LF. Today, most software applications recognize both options, and are able to deal with either encoding of line breaks (until rather recently this was not the case, and using the wrong line breaks could lead to unexpected errors). Our impression is that there is a strong tendency in software development to standardize on the simpler ``only LF'' encoding for line breaks, and we suggest that everybody use this encoding whenever possible.

A code unit is a sequence of bits used to encode character in a encoding. The issue with code units stems from the question how to separate a stream of binary ones and zero, i.e.~bits, into chunks representing Unicode characters. Depending on different use cases, the Unicode Standard offers three different approaches, called UTF-32, UTF-16 and UTF-8.\footnote{The letters UTF stand for either ``Unicode Transformation Format'' or ``Universal Coded Character Set Transformation Format'', but the notion of `transformation' is a legacy notion that does not have meaning anymore. Nevertheless, the designation ``UTF'' (in capitals) has become an official standard designation, but should probably best be read as simply ``Unicode Format''.} The details of this issue is extensively explained in section 2.5 of the Unicode core specification. Basically, \textsc{UTF-32} encodes each character in 32 bits (32 `binary units', i.e.~32 zeros or ones) and is the most disk-space-consuming variant of the three. However, it is the most efficient encoding processing-wise, because the computer simply has to separate each character after 32 bits. In contrast, \textsc{UTF-16} uses only 16 bits per character, which is sufficient for the large majority of Unicode characters, but not for all of them. A special system of `surrogates' is defined within the Unicode Standard to deal with these additional characters. The effect is a more disk-space efficient encoding (approximately half the size), while adding a limited computational overhead. Finally, \textsc{UTF-8} is a more complex system that dynamically encodes each character with the minimally necessary number of bits, choosing either 8, 16 or 32 bits depending on the character. This represents again a strong reduction in space (particularly due to the high frequency of data using ASCII characters that need only 8 bits) at the expense of even more computation necessary to process such strings. However, because of the ever growing computational power of modern machines, the processing overhead is in most practical situations a non-issue, while saving on space is still useful, particularly for sending texts over the Internet. As an effect, UTF-8 has become the dominant encoding on the World Wide Web. We suggest that everybody uses UTF-8 as their default encoding.

A related problem is a general issue about how to store information in computer memory, which is known as \textsc{endianness}. The details of this issue go beyond the scope of this paper. It suffices to realize that there is a difference between \textsc{big-endian} (BE) storage and \textsc{little-endian} (LE) storage. The Unicode Standard offers a possibility to explicitly indicate what kind of storage is used by starting a file with a so-called `Byte Order Mark' (BOM). However, the Unicode Standard does not require the usage of BOM, preferring other non-Unicode methods to signal to computers which kind of endianness is used. This issue only arises with UTF-32 and UTF-16 encodings, and when using the preferred UTF-8, using a BOM is theoretically possible, but strongly dispreferred according to the Unicode Standard. We suggest that everyone tries to prevent the inclusion of BOM in your data.

Together with the topic of normalisation (NFC vs.~NFD, see Pitfall 5 above), this leads to the following nicely cryptic file-encoding recommendation, namely to preferably use \textsc{UTF-8 in NFC without BOM with LF line breaks}, whenever possible.