### 2015-06-05 chapters/pitfalls.tex

\chapter{Unicode pitfalls}
\label{unicode-pitfalls}

% ==========================
\section{Wrong it is not}
\label{wrong-it-is-not}
% ==========================

In this chapter we describe some of the most common pitfalls that we have
encountered when using the Unicode Standard in our own work, or in discussion
with other linguists. This section is not meant as a criticism of the decisions
made by the Unicode Consortium; on the contrary, we aim to highlight where the
technological aspects of the Unicode Standard diverge from many users'
intuitions. What have sometimes been referred to as problems or inconsistencies
in the Unicode Standard are mostly due to legacy compatibility issues, which can
lead to unexpected behavior by linguists using the standard. However, there are
also some cases in which the Unicode Standard has made decisions that
theoretically could have been taken differently, but for some reason or another
(mostly very good reasons) were accepted as they are now. We call behavior that
executes without error but does something different than the user
expected---often unknowingly---a \textsc{pitfall}.

In this context, it is important to realize that the Unicode Standard was not
developed to solve linguistic problems per se, but to offer a consistent
computational environment for written language. In those cases in which the
Unicode Standard behaves differently as expected, we think it is important not
to dismiss Unicode as ``wrong'' or ``deficient'', because our
experience is that in almost all cases the behavior of the Unicode Standard has
been particularly well thought through. The Unicode Consortium has a more
wide-ranging view of matters and often examines important practical use-cases
that from a linguistic point of view are normally not considered. Our general
guideline for dealing with the Unicode Standard is to accept it as it is, and
not to battle windmills. Alternatively, of course, it is possible to actively
engage in the development of the standard itself, an effort that is highly
appreciated by the Unicode Consortium.

% ==========================
\section{Pitfall: Blocks}
\label{pitfall-blocks}
% ==========================

The Unicode code space is subdivided into blocks of contiguous code points. For
example, the block called \textsc{Cyrillic} runs from \uni{0400} till
\uni{04FF}. These blocks arose as an attempt at ordering the enormous amount of
characters in Unicode, but the ideas of blocks very quickly ran into problems.
First, the size of a block is fixed, so when a block is full, a new block will
have to be instantiated somewhere further in the code space. For example, this
led to the blocks \textsc{Cyrillic Supplement}, \textsc{Cyrillic Extended-A}
(both of which are also already full) and \textsc{Cyrillic Extended-B}. Second,
when a specific character already exists, then it is not duplicated in another
block, although the name of the block might indicate that a specific symbol
should be available there. In general, names of blocks are just an approximate
indication of the kind of characters that will be in the block.

The problem with blocks arises because finding the right character among the
thousands of Unicode characters is not easy. Many software applications present
blocks as a primary search mechanism, because the block names suggest where to
look for a particular character. However, when a user searches for an IPA
character in the block \textsc{IPA Extensions}, then many IPA characters will not
be found there. For example, the velar nasal <ŋ> is not part of the block
\textsc{IPA Extensions} because it was already included as \textsc{latin small letter
eng} at \uni{014B} in the block \textsc{Latin Extensions-A}.

In general, finding a specific character in the Unicode Standard is often not
trivial. The names of the blocks can help, but it is not (and never was supposed
to be) a foolproof structure. It is not the goal nor aim of the Unicode
Consortium to provide a user interface to the Unicode Standard. If one often
encounters the problem of needing to find a suitable character, there are
various other useful services for end-users available.\footnote{The Unicode
website offers a basic interface to the code charts at
\url{http://www.unicode.org/charts/index.html}. As a more flexible interface, we
particularly like PopChar from Macility, available for both Macintosh and
Windows. However, there are also various free websites that offer search interfaces
to the Unicode code tables, like \url{http://unicode-search.net} or
\url{http://unicode-search.net}. A further useful approach for searching characters
using shape matching is \url{http://shapecatcher.com}.}

% ==========================
\section{Pitfall: Missing glyphs}
\label{pitfall-missing-glyphs}
% ==========================

The Unicode Standard is often praised (and deservedly so) for solving many of
the perennial problems with the interchange and display of the world's writing
systems. However, a common complaint from users is that, while the praise may be
true, they mostly just see some boxes on their screen instead of those promised
symbols. The problem of course is that users' computers do not have any glyphs
installed matching the Unicode code points in the file they are trying to
inspect. It is important to realize that internally in the computer everything
still works as expected: any handling of Unicode code points works independently
of how they are displayed on the screen. So, although a user might only see
boxes being displayed, this user should be assured that everything is still in
order.

The central problem behind the missing glyphs is that designing actual glyphs
includes a lot of different considerations and it is a time-consuming process.
Many traditional expectations of how specific characters should look like have
to be taken into account when designing glyphs. Those expectations are often not
well documented, and it is mostly up to the knowledge and experience of the font
designer to try and conform to them as good as possible. Therefore, most
designers produce fonts only including glyphs for certain parts of the Unicode
Standard, namely for those characters they feel comfortable with. At the same
time, the number of characters defined by the Unicode Standard is growing with
each new version, so it is neigh impossible for any designer to produce glyphs
for all characters. The result of this is that, almost necessarily, each font
only includes glyphs for a subset of the characters in the Unicode Standard.

The simple solution to missing glyphs is thus to install additional fonts
providing additional glyphs. For the more exotic characters there is often not
much choice. There are a few particularly large fonts that might be considered.
First, there is the \textsc{Everson Mono} font made by Michael Everson, which
currently includes 9,756 different glyphs (not including Chinese) updated up to
Unicode 7.0.\footnote{Everson Mono is available as shareware at
\url{http://www.evertype.com/emono/}.} Already a bit older is the \textsc{Titus
Cyberbit Basic} font made by Jost Gippert and Carl-Martin Bunz, which includes
10,044 different glyphs (not including Chinese), but not including newer
characters added after Unicode 4.0.\footnote{Titus Cyberbit Basic is available
at \url{http://titus.fkidg1.uni-frankfurt.de/unicode/tituut.asp}.}

Further, we suggest to always install at least one so-called \textsc{fall-back
font}, which provides glyphs that at least show the user some information about
the underlying encoded character. Apple Macintoshes have such a font (which is
invisible to the user), which is designed by Michael Everson and made available
for other systems through the Unicode Consortium.\footnote{The Apple/Everson
fallback font is available for non-Macintosh users at \newline
\url{http://www.unicode.org/policies/lastresortfont\_eula.html}.} Further, the
\textsc{GNU Unifont} is a clever way to produce bitmaps approximating the
intended glyph of each available character, updated to Unicode 7.0.\footnote{The
GNU Unifont is available at \url{http://unifoundry.com/unifont.html}.} Finally,
the Summer Institute of Linguistics provides a \textsc{SIL Unicode BMP Fallback
Font}, currently available up to Unicode version 6.1. This font does not even
attempt to show a real glyph, but only shows the hexadecimal code inside a box
for each character, so a user can at least see the Unicode codepoint of the
character to be displayed.\footnote{The SIL Unicode BMP Fallback Font is
available at \newline \url{http://scripts.sil.org/UnicodeBMPFallbackFont}.}

% ==========================
\section{Pitfall: Faulty rendering}
\label{pitfall-faulty-rendering}
% ==========================

A similar complaint to missing glyphs, discussed previously, is that while there
might be a glyph being displayed, it does not look right. There are two
reasons for unexpected visual display, namely automatic font substitution and
faulty rendering. Like missing glyphs, any such problems are independent from
the Unicode Standard. The Unicode Standard only includes very general
information about characters and leaves the specific visual display to others to
decide on. Any faulty display is thus not to be blamed on the Unicode
Consortium, but on a complex interplay of different mechanisms happening in a
computer to turn Unicode codepoints into visual symbols. We will only sketch a
few aspects of this complex interplay here.

Most modern software applications (like Microsoft Word) offer some approach to
\textsc{automatic font substitution}. This means that when a text is written in
a specific font (e.g.~Times New Roman) and an inserted Unicode character does not
have a glyph within this font, then the software application will automatically
search for another font to display the glyph. The result will be that this
specific glyph will look slightly different from the others. This mechanism
works differently depending on the software application, and mostly only limited
user influence is expected and little feedback is given, which might be rather
frustrating to font-aware users.\footnote{For example, Apple Pages does not give
any feedback that a font is being replaced, and the user does not seem to have
any influence on the choice of replacement (except by manually marking all
occurrences). In contrast, Microsoft Word does indicate the font replacement by
showing the name in the font menu of the font replacement. However, Word simply
changes the font completely, so any text written after the replacement is written in a
different font as before. Both behaviors leave much to be desired.}

The other problem with visual display is related to the so-called \textsc{font
rendering}. Font rendering refers to the process of the actual positioning of
Unicode characters on a page of written text. This positioning is actually a
highly complex problem, and many things can go wrong in the process. Well-known
rendering problems, like proportional glyph size or ligatures are reasonably
well understood. However, even within the Latin script, the positioning of
diacritics relative to a base character is still a widespread problem.
Especially when more than one diacritic is supposed to be placed above (or
below) each other, this often leads to unexpected effects in many modern
software applications. The problems arising in Arabic and in many southeast
Asian scripts (like Devanagari or Burmese) are even much more complex. There are
basically three different approaches to font rendering. The most widespread is
Adobe's and Microsoft's \emph{OpenType }system. This approach makes it
relatively easy for font developers, as the font itself does not include all
details about the precise placement of individual characters. For those details,
additional script-descriptions are necessary. All of those systems can lead to
unexpected behavior.\footnote{For more details about OpenType, see
\url{http://www.adobe.com/products/type/opentype.html} and
\url{http://www.microsoft.com/typography/otspec/}. Additional systems for
complex text layout are, among others, Microsoft's DirectWrite
\url{https://msdn.microsoft.com/library/dd368038.aspx} and the open-source
project HarfBuzz \url{http://www.freedesktop.org/wiki/Software/HarfBuzz/}.}
Alternative systems are Apple's \emph{Apple Advanced Typography} (AAT) and the
open-source \emph{Graphite} system from the Summer Institute of Linguistics
(SIL).\footnote{More information about AAT can be found at
\url{https://developer.apple.com/fonts/}. SIL's Graphite is described in detail
at
\url{http://scripts.sil.org/cms/scripts/page.php?site\_id=projects\&item\_id=graphite\_home}.}
In both of these systems, a larger burden is placed on the description inside
the font.

% ==========================
\section{Pitfall: Homoglyphs}
\label{pitfall-homoglyphs}
% ==========================

Homoglyphs are visually indistinguishable glyphs (or highly similar glyphs) that
have different code points in the Unicode Standard and thus different character
semantics. As a principle, the Unicode Standard does not specify how a character
appears visually on the page or the screen. So in most cases, a different
appearance is caused by the specific design of a font, or by user-settings like
size or boldface. Taking an example already discussed in
Section~\ref{pitfall-homoglyphs}, the following symbols <g {\large \textit{g}}
\textbf{g} {\fontspec{ArialMT} {\small g} \textit{g} \textbf{g}}> are different
glyphs of the same character, i.e.~they may be rendered differently depending on
the typography being used, but they all share the same code point (viz.
\textsc{latin small letter g} at \uni{0067}). In contrast, the symbols
<{\fontspec{EversonMono}AАΑᎪᗅᴀꓮ𐊠𐌀𝖠𝙰}> are all different code points,
although they look highly similar---in some cases even sharing exactly the same
glyph in some fonts. All these different A-like characters include the following
code points in the Unicode Standard:

\begin{itemize}
	\item <{\fontspec{EversonMono}A}> \textsc{latin capital letter a}, at \uni{0041} 
	\item <{\fontspec{EversonMono}А}> \textsc{cyrillic capital letter a}, at \uni{0410} 
	\item <{\fontspec{EversonMono}Α}> \textsc{greek capital letter alpha}, at \uni{0391} 
	\item <{\fontspec{EversonMono}Ꭺ}> \textsc{cherokee letter go}, at \uni{13AA} 
	\item <{\fontspec{EversonMono}ᗅ}> \textsc{canadian syllabics carrier gho}, at \uni{15C5} 
	\item <{\fontspec{EversonMono}ᴀ}> \textsc{latin small letter capital a}, at \uni{1D00} 
	\item <{\fontspec{EversonMono}ꓮ}> \textsc{lisu letter a}, at \uni{A4EE} 
%	\item <{\fontspec{EversonMono}Ａ}> \textsc{fullwidth latin capital letter a}, at \uni{FF21} 
	\item <{\fontspec{EversonMono}𐊠}> \textsc{carian letter a}, at \uni{102A0} 
	\item <{\fontspec{EversonMono}𐌀}> \textsc{old italic letter a}, at \uni{10300} 
	\item <{\fontspec{EversonMono}𝖠}> \textsc{mathematical sans-serif capital a}, \uni{1D5A0} 
	\item <{\fontspec{EversonMono}𝙰}> \textsc{mathematical monospace capital a}, at \uni{1D670} 
\end{itemize}

The existence of such homoglyphs is partly due to legacy compatibility, but for
the most part these characters are simply different characters that happen to
look similar.\footnote{A particularly nice interface to look for homoglyphs is
\url{http://shapecatcher.com}, based on the principle of recognizing shapes
\citep{Belongie2002}.} Yet, they are supposed to behave different from the
perspective of a font designer. For example, when designing a Cyrillic font, the
<A> will have different aesthetics and different traditional expectation
compared to a Latin <A>.

Such homoglyphs are a widespread problem for consistent encoding. Although for
most users it looks like the words <voces> and <νοсеѕ> are identical, in actual
fact they do not even share a single code point.\footnote{The first words
consists completely of Latin characters, namely \unif{0076}, \unif{006F},
\unif{0063}, \unif{0065} and \unif{0073}, while the second is a mix of Cyrillic and
Greek characters, namely \unif{03BD}, \unif{03BF}, \unif{0041}, \unif{0435} and
\unif{0455}.} For computers these two words are completely different entities.
Commonly, when users with Cyrillic or Greek keyboards have to type some
Latin-based orthography, they mix similar looking Cyrillic or Greek characters
into their text, because those characters are so much easier to type. Similarly,
when users want to enter an unusual symbol, they normally search by visual
impression in their favorite software application, and just pick something that
looks reasonably alike to what they expect the glyph to look like.

% ==========================
\section{Pitfall: Canonical equivalence}
\label{pitfall-canonical-equivalence}
% ==========================

For some characters, there is more than one possible encoding in the Unicode
Standard. This is a possible pitfall, as this would mean that for the computer
there exist multiple different entities that for a user are the same. This
would, for example, lead to problems with searching, as the computer would
search for specific encodings, and not find all expected characters. As a
solution, the Unicode Standard includes a notion of \textsc{canonical
equivalence}. Different encodings are explicitly declared as equivalent in the
Unicode Standard code tables. Further, to harmonize all encodings in a specific
piece of text, the Unicode Standard proposes a mechanism of
\textsc{normalization}.

Consider for example the characters and following Unicode code points:
\begin{itemize}
	\def\labelenumi{\arabic{enumi}.} 
	\item <Å> \textsc{latin capital letter a with ring above} \uni{00C5} 
	\item <Å> \textsc{angstrom sign} \uni{212B}
	\item <Å> \textsc{latin capital letter a} \uni{0041}
	+ \textsc{combining ring above} \uni{030A}
\end{itemize}

The character, represented here by glyph <Å>, is encoded in the Unicode Standard
in the first two examples by a single-character sequence; each is assigned a
different code point. In the third example, the glyph is encoded in a
multiple-character sequence that is composed of two character code points. All
three sequences are \textsc{canonically equivalent}, i.e.~they are strings that
represent the same abstract character and because they are not distinguishable
by the user, the Unicode Standard requires them to be treated the same in
regards to their behavior and appearance. Nevertheless, they are encoded
differently. For example, if one were to search an electronic text (with
software that does not apply Unicode Standard normalization) for
\textsc{angstrom sign} (\uni{212B}), instances of \textsc{latin capital letter a
with ring above} (\uni{00C5}) would not be found.

In other words, there are equivalent sequences of Unicode characters that should
be normalized, i.e.~transformed into a unique Unicode-sanctioned representation
of a character sequence called a \textsc{normalization form}. Unicode provides a
Unicode Normalization Algorithm, which essentially puts combining marks
into a specific logical order and it defines decomposition and composition
transformation rules to convert each string into one of four normalization
forms. We will discuss here the two most relevant normalization forms: NFC and
NFD.\@

The first of the three characters above is considered the \textsc{Normalization
Form C (NFC)}, where \textsc{C} stands for composition. When the process of NFC
normalization is applied to the character sequences in 2 and 3, both sequences
are normalized into the \textsc{pre-composed} character sequence in 1. Thus all
three canonical character sequences are standardized into one composition form
in NFC.\@The other central Unicode normalization form is the
\textsc{Normalization Form D (NFD)}, where \textsc{D} stands for decomposition.
When NFD is applied to the three examples above, all three, including
importantly the single-character sequences in 1 and 2, are normalized into the
\textsc{decomposed} multiple-sequence of characters in 3. Again, all three are
then logically equivalent and therefore comparable and syntactically
interoperable.

As illustrated, some characters in the Unicode Standard have alternative
representations (in fact, many do), but the Unicode Normalization Algorithm can
be used to transform certain sequences of characters into canonical (or
compatible) forms to test for equivalency. To determine equivalence, each
character in the Unicode Standard is associated with a combining class, which is
formally defined as a character property called \textsc{canonical combining
class} which is specified in the Unicode Character Database. The combining class
assigned to each code point is a numeric value between 0 and 254 and is used by
the Unicode Canonical Ordering Algorithm to determine which sequences of
combining marks are canonically equivalent. Normalization forms, as very briefly
described above, can be used to ensure character equivalence by ordering
character sequences so that they can be faithfully compared.

It is very important to note that any software applications that is Unicode
Standard compliant is free to change the character stream from one
representation to another. This means that a software application may compose,
decompose or reorder characters as its developers desire; as long as the
resultant strings are canonically equivalent to the original. This might lead to
unexpected behavior for users. Various players, like the Unicode Consortium, the
W{\large 3}C, or the TEI recommend NFC in most user-directed situations, and some
software applications that we tested indeed seem to automatically convert
strings into NFC.\footnote{See the summary of various recommendation here:
\url{http://www.win.tue.nl/~aeb/linux/uc/nfc_vs_nfd.html}.}
This means in practice that if a user, for example, enters <a> and < ̀>,
i.e.~\textsc{latin small letter a} at \uni{0061} and \textsc{combining grave
accent} at \uni{0300}, this might be converted into <à>, i.e.~\textsc{latin
small letter a with grave} at \uni{00E0}.\footnote{The behavior of software
applications can be quite erratic in this respect. For example, Apple's TextEdit
does not do any conversion on text entry. However, when you copy and paste some
text inside the same document in rich text mode (i.e.~RTF-format), it
will be transformed into NFC on paste. Saving a document does not do any
conversion to the glyphs on screen, but it will save the characters in NFC.}

% ==========================
\section{Pitfall: Characters are not glyphs}
\label{pitfall-characters-are-not-glyphs}
% ==========================

A central principle of Unicode is the distinction between character and glyph. A
character is the abstract notion of a symbol in a writing system, while a glyph
is the concrete drawing of such a symbol. In practice, there is a complex
interaction between characters and glyphs. A single Unicode character may of
course be rendered as a single glyph. However, a character may also be a
\textit{piece} of a glyph, or vice-versa. Actually, all possible relations
between glyphs and characters are attested.

First, \textit{a single character may have different contextually determined
glyphs.} For example, characters in writing systems like Hebrew and Arabic have
different glyphs depending on where they appear in a word. Some letters in
Hebrew change their form at the end of the word, and in Arabic, primary letters
have four contextually-sensitive variants (isolated, word initial, medial and
final). Second, \textit{a single character may be rendered as a sequence of
multiple glyphs.} For example, in Tamil one Unicode character may result in a
combination of a consonant and vowel, which are rendered as two adjacent glyphs
by fonts that supports Tamil. Third, \textit{a single glyph may be a combination
of multiple characters.} For example, the ligature <ï¬>, a single glyph, is the
result of two characters, <f> and <i>, that have undergone glyph substitution by
font rendering. Like contextually-determined glyphs, ligatures are (intended)
artifacts of text processing instructions. Finally, \textit{a single glyph may
be a part of a character} (e.g.~diacritics).

Further, the rendering of a glyph is also always dependent on the font being
used. For example, the Unicode character \textsc{latin small letter g} appears
as and in the Computer Modern and Courier fonts because their typefaces are
designed differently. Furthermore, font face may change the visual appearance of
a character, for example Times New Roman two-story <a> changes to a single-story
a in italics <\emph{a}>. This becomes a real problem for some phonetic
typesetting (see Section~\ref{pitfall-missing-glyphs}).

In sum, character-to-glyph mappings are complex technological issues that the
Unicode Consortium has had to address in the development of the Unicode
Standard, but for the lay user they can be utterly confusing because visual
rendering does not indicate logical encoding, or, in other words, what you see
isn't (necessarily) what you get.

% ==========================
\section{Pitfall: File formats}
\label{pitfall-file-formats}
% ==========================

Unicode is a character encoding standard, but these characters of course
actually appear inside some kind of file. The most basic Unicode-based file
format is pure line-based text, i.e.~strings of Unicode-encoded characters
separated by line breaks (note that these line breaks are what for most people
intuitively corresponds to paragraph breaks). Unfortunately, even within this
apparently basic setting there exist a multitude of variants. In general, these
different possibilities are well-understood in the software industry, and
nowadays they normally do not lead to any problems for the end user. However,
there are some situations in which a user is suddenly confronted with cryptic
questions in the user interface involving abbreviations like \textsc{lf},
\textsc{cr}, \textsc{be}, \textsc{le} or \textsc{bom}. Most prominently this
occurs with exporting or importing data in several software applications from
Microsoft. Basically, there are two different issues involved. First, the
encoding of line breaks and, second, the encoding of the Unicode characters into
\textsc{code units} and the related issue of \textsc{endianness}.

The issue with \textsc{line breaks} originated with the instructions necessary
to direct a printing head to a new line. This involves two movements, known as
\textsc{carriage return (cr)} (returning the printing head to the start of the
line on the page) and \textsc{line feed (lf)} (moving the printing head to the
next line on the page). Physically, these are two different events, but
conceptually together they form one action. In the history of computing, various
encodings of line breaks have been used (e.g.~\textsc{cr+lf}, \textsc{lf+cr},
only \textsc{lf}, or only \textsc{cr}). Currently, all Unix and Unix-derived
systems use only \textsc{lf} as code for a line break, while software from
Microsoft still uses a combination of \textsc{cr} and \textsc{lf}. Today, most
software applications recognize both options, and are able to deal with either
encoding of line breaks (until rather recently this was not the case, and using
the wrong line breaks could lead to unexpected errors). Our impression is that
there is a strong tendency in software development to standardize on the simpler
\textit{only \textsc{lf}} encoding for line breaks, and we suggest that
everybody use this encoding whenever possible.

A code unit is a sequence of bits used to encode character in a encoding. The
issue with code units stems from the question how to separate a stream of binary
ones and zero, i.e.~bits, into chunks representing Unicode characters. Depending
on different use cases, the Unicode Standard offers three different approaches,
called UTF-32, UTF-16 and UTF-8.\footnote{The letters UTF stand for either
\textsc{Unicode Transformation Format} or \textsc{Universal Coded Character Set
Transformation Format}, but the notion of \textit{transformation} is a legacy
notion that does not have meaning anymore. Nevertheless, the designation
\textsc{UTF} (in capitals) has become an official standard designation, but
should probably best be read as simply \textit{Unicode Format}.} The details of
this issue is extensively explained in section 2.5 of the Unicode core
specification. Basically, \textsc{UTF-32} encodes each character in 32 bits (32
\textit{bi}nary un\textit{ts}, i.e.~32 zeros or ones) and is the most
disk-space-consuming variant of the three. However, it is the most efficient
encoding processing-wise, because the computer simply has to separate each
character after 32 bits. In contrast, \textsc{UTF-16} uses only 16 bits per
character, which is sufficient for the large majority of Unicode characters, but
not for all of them. A special system of `surrogates' is defined within the
Unicode Standard to deal with these additional characters. The effect is a more
disk-space efficient encoding (approximately half the size), while adding a
limited computational overhead. Finally, \textsc{UTF-8} is a more complex system
that dynamically encodes each character with the minimally necessary number of
bits, choosing either 8, 16 or 32 bits depending on the character. This
represents again a strong reduction in space (particularly due to the high
frequency of data using ASCII characters that need only 8 bits) at the expense
of even more computation necessary to process such strings. However, because of
the ever growing computational power of modern machines, the processing overhead
is in most practical situations a non-issue, while saving on space is still
useful, particularly for sending texts over the Internet. As an effect, UTF-8
has become the dominant encoding on the World Wide Web. We suggest that
everybody uses UTF-8 as their default encoding.

A related problem is a general issue about how to store information in computer
memory, which is known as \textsc{endianness}. The details of this issue go
beyond the scope of this paper. It suffices to realize that there is a
difference between \textsc{big-endian} (BE) storage and \textsc{little-endian}
(LE) storage. The Unicode Standard offers a possibility to explicitly indicate
what kind of storage is used by starting a file with a so-called `Byte Order
Mark' (BOM). However, the Unicode Standard does not require the usage of BOM,
preferring other non-Unicode methods to signal to computers which kind of
endianness is used. This issue only arises with UTF-32 and UTF-16 encodings, and
when using the preferred UTF-8, using a BOM is theoretically possible, but
strongly dispreferred according to the Unicode Standard. We suggest that
everyone tries to prevent the inclusion of BOM in your data.

Together with the topic of normalization (NFC vs.~NFD, see Pitfall 5 above),
this leads to the following nicely cryptic file-encoding recommendation, namely
to preferably use \textsc{UTF-8 in NFC without BOM with LF line breaks},
whenever possible.