### 2015-06-05 localpackages.tex

% add all extra packages you need to load to this file

\usepackage{amsmath} 
\usepackage{unicode-math}
\setmathfont{Asana-Math.otf}
\usepackage[libertine]{newtxmath}

\usepackage{enumitem}
\setitemize{noitemsep}
\setenumerate{noitemsep}

\usepackage{tabularx}

\widowpenalty=10000
\clubpenalty=10000

% I have tried to use monospaced numbers in the TOC, but this does not work...
\settocstylefeature{\fontspec[Numbers=Monospaced]{LinLibertineO}}
%\settocfeature{\fontspec[Numbers=Monospaced]{LinLibertineO}}

% other ideas that also don't give the desired results
%\usepackage{tocloft}
%\renewcommand{\cftXpagefont}{\fontspec[Numbers=Monospaced]{LinLibertineO}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                                              %%%
%%%           Examples                           %%%
%%%                                              %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\usepackage{lsp-gb4e}

%% to add additional information to the right of examples, uncomment the following line
% \usepackage{jambox}
%% if you want the source line of examples to be in italics, uncomment the following line
% \renewcommand{\exfont}{\itshape}

\usepackage{listings}

\lstset{ %
  backgroundcolor=\color{white},   % choose the background color; you must add \usepackage{color} or \usepackage{xcolor}
  basicstyle=\footnotesize\ttfamily,        % the size of the fonts that are used for the code 
  keywordstyle=\color{blue!60!black},       % keyword style
  language=XML,                 % the language of the code 
  stringstyle=\color{green!60!black},     % string literal style 
  morekeywords={token,xlink:href, Action, Value, Cursor,LogEvent}
}