### 2015-06-18 chapters/tab:unicode_diacritics.tex

\tablecaption{Strict-IPA diacritics with Unicode encodings}\label{tab:ipa_diacritics}
 \tablefirsthead{
   \toprule
   & Code & Unicode name & IPA name \\ 
   \midrule
   }
 \tablehead{
   \multicolumn{4}{c}{
     \captionsize\tablename\ \thetable{} 
     Strict-IPA diacritics with Unicode encodings --- \textit{continued}
     } \\
   \toprule
   & Code & Unicode name & IPA name \\ 
   \midrule
   }
 \tabletail{
   \bottomrule
   \multicolumn{4}{r}{
     \captionsize\textit{continued on next page}
     } \\
   }
 \tablelasttail{\bottomrule}

\begin{center}
\begin{xtabular}{ l L{1.2cm} L{5.6cm} L{4cm} }
{\large \ ̼} & \unif{033C} & \textsc{combining seagull below} & linguolabial \\
{\large \ ̪} & \unif{032A} & \textsc{combining bridge below} & dental \\
{\large \ ̻} & \unif{033B} & \textsc{combining square below} & laminal \\
{\large \ ̺} & \unif{033A} & \textsc{combining inverted bridge below} & apical \\
{\large \ ̟} & \unif{031F} & \textsc{combining plus sign below} & advanced \\
{\large \ ̠} & \unif{0320} & \textsc{combining minus sign below} & retracted \\
{\large \ ̝} & \unif{031D} & \textsc{combining up tack below} & raised \\
{\large \ ̞} & \unif{031E} & \textsc{combining down tack below} & lowered \\
{\large \ ̘} & \unif{0318} & \textsc{combining left tack below} & advanced tongue root \\
{\large \ ̙} & \unif{0319} & \textsc{combining right tack below} & retracted tongue root \\
{\large \ ̜} & \unif{031C} & \textsc{combining left half ring below} & less rounded \\
{\large \ ̹} & \unif{0339} & \textsc{combining right half ring below} & more rounded \\
{\large \ ̬} & \unif{032C} & \textsc{combining caron below} & voiced \\
{\large \ ̥} & \unif{0325} & \textsc{combining ring below} & voiceless \\
{\large \ ̰} & \unif{0330} & \textsc{combining tilde below} & creaky voiced \\
{\large \ ̤} & \unif{0324} & \textsc{combining diaeresis below} & breathy voiced \\
{\large \ ̩} & \unif{0329} & \textsc{combining vertical line below} & syllabic \\
{\large \ ̯} & \unif{032F} & \textsc{combining inverted breve below} & non-syllabic \\
\midrule
{\large \ ̃} & \unif{0303} & \textsc{combining tilde} & nasalized \\
{\large \ ̈} & \unif{0308} & \textsc{combining diaeresis} & centralized \\
{\large \ ̽} & \unif{033D} & \textsc{combining x above} & mid-centralized \\
{\large \ ̆} & \unif{0306} & \textsc{combining breve} & extra-short \\
{\large ̚} & \unif{031A} & \textsc{combining left angle above} & no audible release \\
\midrule
{\large ˞} & \unif{02DE} & \textsc{modifier letter rhotic hook} & rhotacized \\
{\large ˡ} & \unif{02E1} & \textsc{modifier letter small l} & lateral release \\
{\large ⁿ} & \unif{207F} & \textsc{superscript latin small letter n} & nasal release \\
{\large ʷ} & \unif{02B7} & \textsc{modifier letter small w} & labialized \\
{\large ʲ} & \unif{02B2} & \textsc{modifier letter small j} & palatalized \\
{\large ˠ} & \unif{02E0} & \textsc{modifier letter small gamma} & velarized \\
{\large ˤ} & \unif{02E4} & \textsc{modifier letter small reversed glottal stop} & pharyngealized \\
{\large ʰ} & \unif{02B0} & \textsc{modifier letter small h} & aspirated \\
{\large ʼ} & \unif{02BC} & \textsc{modifier letter apostrophe} & ejective \\
{\large ː} & \unif{02D0} & \textsc{modifier letter triangular colon} & long \\
{\large ˑ} & \unif{02D1} & \textsc{modifier letter half triangular colon} & half-long \\
\midrule
{˥} & \unif{02E5} & \textsc{modifier letter extra-high tone bar} & extra high tone \\
{˦} & \unif{02E6} & \textsc{modifier letter high tone bar} & high tone \\
{˧} & \unif{02E7} & \textsc{modifier letter mid tone bar} & mid tone \\
{˨} & \unif{02E8} & \textsc{modifier letter low tone bar} & low tone \\
{˩} & \unif{02E9} & \textsc{modifier letter extra-low tone bar} & extra low tone \\
{↑} & \unif{2191} & \textsc{upwards arrow} & upstep \\
{↓} & \unif{2193} & \textsc{downwards arrow} & downstep \\
{↗} & \unif{2197} & \textsc{north east arrow} & global rise \\
{↘} & \unif{2198} & \textsc{south east arrow} & global fall \\
\midrule
{\large ˈ} & \unif{02C8} & \textsc{modifier letter vertical line} & primary stress \\
{\large ˌ} & \unif{02CC} & \textsc{modifier letter low vertical line} & secondary stress \\
{\large.} & \unif{002E} & \textsc{full stop} & syllable break \\
{|} & \unif{007C} & \textsc{vertical line} & minor group break (foot) \\
{‖} & \unif{2016} & \textsc{double vertical line} & major group break (intonation) \\
{\fontspec{CharisSIL} ‿} & \unif{203F} & \textsc{undertie} & linking (absence of a break) \\
\end{xtabular}
\end{center}