### 2015-06-18 chapters/tab:unicode_letters.tex

\tablecaption{Strict-IPA letters with Unicode encodings}\label{tab:ipa_letters}
 \tablefirsthead{
   \toprule
   & Code & Unicode name & IPA name \\ 
   \midrule
   }
 \tablehead{
   \multicolumn{4}{c}{
     \captionsize\tablename\ \thetable{} 
     Strict-IPA letters with Unicode encodings --- \textit{continued}
     } \\
   \toprule
   & Code & Unicode name & IPA name \\ 
   \midrule
   }
 \tabletail{
   \bottomrule
   \multicolumn{4}{r}{
     \captionsize\textit{continued on next page}
     } \\
   }
 \tablelasttail{\bottomrule}

\begin{center}
\begin{xtabular}{ l L{1.2cm} L{5.4cm} L{4.2cm} }
a & \unif{0061} & \textsc{latin small letter a} & open front unrounded \\ 
æ & \unif{00E6} & \textsc{latin small letter ae} & raised open front unrounded \\ 
ɐ & \unif{0250} & \textsc{latin small letter turned a} & lowered schwa \\ 
ɑ & \unif{0251} & \textsc{latin small letter alpha} & open back unrounded \\ 
ɒ & \unif{0252} & \textsc{latin small letter turned alpha} & open back rounded \\ 
b & \unif{0062} & \textsc{latin small letter b} & voiced bilabial plosive \\ 
ʙ & \unif{0299} & \textsc{latin letter small capital b} & voiced bilabial trill \\ 
ɓ & \unif{0253} & \textsc{latin small letter b with hook} & voiced bilabial implosive \\ 
c & \unif{0063} & \textsc{latin small letter c} & voiceless palatal plosive \\ 
ç & \unif{00E7} & \textsc{latin small letter c with cedilla} & voiceless palatal fricative \\ 
ɕ & \unif{0255} & \textsc{latin small letter c with curl} & voiceless alveolo-palatal fricative \\ 
d & \unif{0064} & \textsc{latin small letter d} & voiced alveolar plosive \\ 
ð & \unif{00F0} & \textsc{latin small letter eth} & voiced dental fricative \\ 
ɖ & \unif{0256} & \textsc{latin small letter d with tail} & voiced retroflex plosive \\ 
ɗ & \unif{0257} & \textsc{latin small letter d with hook} & voiced dental/alveolar implosive \\ 
% \charis{ᶑ} & \unif{1D91} & \textsc{latin small letter d with hook and tail} & voiced retroflex implosive \\ 
e & \unif{0065} & \textsc{latin small letter e} & close-mid front unrounded \\ 
ə & \unif{0259} & \textsc{latin small letter schwa} & mid-central schwa \\ 
ɛ & \unif{025B} & \textsc{latin small letter open e} & open-mid front unrounded \\ 
ɘ & \unif{0258} & \textsc{latin small letter reversed e} & close-mid central unrounded \\ 
ɜ & \unif{025C} & \textsc{latin small letter reversed open e} & open-mid central unrounded \\ 
ɞ & \unif{025E} & \textsc{latin small letter closed reversed open e} & open-mid central rounded \\ 
f & \unif{0066} & \textsc{latin small letter f} & voiceless labiodental fricative \\ 
ɡ & \unif{0261} & \textsc{latin small letter script g} & voiced velar plosive \\ 
ɢ & \unif{0262} & \textsc{latin letter small capital g} & voiced uvular plosive \\ 
ɠ & \unif{0260} & \textsc{latin small letter g with hook} & voiced velar implosive \\ 
ʛ & \unif{029B} & \textsc{latin letter small capital g with hook} & voiced uvular implosive \\ 
ɤ & \unif{0264} & \textsc{latin small letter rams horn} & close-mid back unrounded \\ 
ɣ & \unif{0263} & \textsc{latin small letter gamma} & voiced velar fricative \\ 
h & \unif{0068} & \textsc{latin small letter h} & voiceless glottal fricative \\ 
ħ & \unif{0127} & \textsc{latin small letter h with stroke} & voiceless pharyngeal fricative \\ 
ʜ & \unif{029C} & \textsc{latin letter small capital h} & voiceless epiglottal fricative \\ 
ɦ & \unif{0266} & \textsc{latin small letter h with hook} & voiced glottal fricative \\ 
ɧ & \unif{0267} & \textsc{latin small letter heng with hook} & simultaneous voiceless postalveolar+velar fricative \\ 
ɥ & \unif{0265} & \textsc{latin small letter turned h} & voiced labial-palatal approximant \\ 
i & \unif{0069} & \textsc{latin small letter i} & close front unrounded \\ 
ɪ & \unif{026A} & \textsc{latin letter small capital i} & lax close front unrounded \\ 
ɨ & \unif{0268} & \textsc{latin small letter i with stroke} & close central unrounded \\ 
j & \unif{006A} & \textsc{latin small letter j} & voiced palatal approximant \\ 
ʝ & \unif{029D} & \textsc{latin small letter j with crossed tail} & voiced palatal fricative \\ 
ɟ & \unif{025F} & \textsc{latin small letter dotless j with stroke} & voiced palatal plosive \\ 
ʄ & \unif{0284} & \textsc{latin small letter dotless j with stroke and hook} & voiced palatal implosive \\ 
k & \unif{006B} & \textsc{latin small letter k} & voiceless velar plosive \\ 
l & \unif{006C} & \textsc{latin small letter l} & voiced alveolar lateral approximant \\ 
ʟ & \unif{029F} & \textsc{latin letter small capital l} & voiced velar lateral approximant \\ 
% ɫ & \unif{026B} & \textsc{latin small letter l with middle tilde} & voiced pharyngealized alveolar lateral approximant \\ 
ɬ & \unif{026C} & \textsc{latin small letter l with belt} & voiceless alveolar lateral fricative \\ 
ɭ & \unif{026D} & \textsc{latin small letter l with retroflex hook} & voiced retroflex lateral approximant \\ 
ɮ & \unif{026E} & \textsc{latin small letter lezh} & voiced alveolar lateral fricative \\ 
ʎ & \unif{028E} & \textsc{latin small letter turned y} & voiced palatal lateral approximant \\ 
m & \unif{006D} & \textsc{latin small letter m} & voiced bilabial nasal \\ 
ɱ & \unif{0271} & \textsc{latin small letter m with hook} & voiced labiodental nasal \\ 
n & \unif{006E} & \textsc{latin small letter n} & voiced alveolar nasal \\ 
ɴ & \unif{0274} & \textsc{latin letter small capital n} & voiced uvular nasal \\ 
ɲ & \unif{0272} & \textsc{latin small letter n with left hook} & voiced palatal nasal \\ 
ɳ & \unif{0273} & \textsc{latin small letter n with retroflex hook} & voiced retroflex nasal \\ 
ŋ & \unif{014B} & \textsc{latin small letter eng} & voiced velar nasal \\ 
o & \unif{006F} & \textsc{latin small letter o} & close-mid back rounded \\ 
ø & \unif{00F8} & \textsc{latin small letter o with stroke} & close-mid front rounded \\ 
œ & \unif{0153} & \textsc{latin small ligature oe} & open-mid front rounded \\ 
ɶ & \unif{0276} & \textsc{latin letter small capital oe} & open front rounded \\ 
ɔ & \unif{0254} & \textsc{latin small letter open o} & open-mid back rounded \\ 
ɵ & \unif{0275} & \textsc{latin small letter barred o} & close-mid central rounded \\ 
p & \unif{0070} & \textsc{latin small letter p} & voiceless bilabial plosive \\ 
ɸ & \unif{0278} & \textsc{latin small letter phi} & voiceless bilabial fricative \\ 
q & \unif{0071} & \textsc{latin small letter q} & voiceless uvular plosive \\ 
r & \unif{0072} & \textsc{latin small letter r} & voiced alveolar trill \\ 
ʀ & \unif{0280} & \textsc{latin letter small capital r} & voiced uvular trill \\ 
ɹ & \unif{0279} & \textsc{latin small letter turned r} & voiced alveolar approximant \\ 
ɺ & \unif{027A} & \textsc{latin small letter turned r with long leg} & voiced alveolar lateral flap \\ 
ɻ & \unif{027B} & \textsc{latin small letter turned r with hook} & voiced retroflex approximant \\ 
ɽ & \unif{027D} & \textsc{latin small letter r with tail} & voiced retroflex tap \\ 
ɾ & \unif{027E} & \textsc{latin small letter r with fishhook} & voiced alveolar tap \\ 
ʁ & \unif{0281} & \textsc{latin letter small capital inverted r} & voiced uvular fricative \\ 
s & \unif{0073} & \textsc{latin small letter s} & voiceless alveolar fricative \\ 
ʂ & \unif{0282} & \textsc{latin small letter s with hook} & voiceless retroflex fricative \\ 
ʃ & \unif{0283} & \textsc{latin small letter esh} & voiceless postalveolar fricative \\ 
t & \unif{0074} & \textsc{latin small letter t} & voiceless alveolar plosive \\ 
ʈ & \unif{0288} & \textsc{latin small letter t with retroflex hook} & voiceless retroflex plosive \\ 
u & \unif{0075} & \textsc{latin small letter u} & close back rounded \\ 
ʉ & \unif{0289} & \textsc{latin small letter u bar} & close central rounded \\ 
ɯ & \unif{026F} & \textsc{latin small letter turned m} & close back unrounded \\ 
ɰ & \unif{0270} & \textsc{latin small letter turned m with long leg} & voiced velar approximant \\ 
ʊ & \unif{028A} & \textsc{latin small letter upsilon} & lax close back rounded \\ 
v & \unif{0076} & \textsc{latin small letter v} & voiced labiodental fricative \\ 
ʋ & \unif{028B} & \textsc{latin small letter v with hook} & voiced labiodental approximant \\ 
\charis{ⱱ} & \unif{2C71} & \textsc{latin small letter v with right hook} & voiced labiodental tap \\ 
ʌ & \unif{028C} & \textsc{latin small letter turned v} & open-mid back unrounded \\ 
w & \unif{0077} & \textsc{latin small letter w} & voiced labial-velar approximant \\ 
ʍ & \unif{028D} & \textsc{latin small letter turned w} & voiceless labial-velar fricative \\ 
x & \unif{0078} & \textsc{latin small letter x} & voiceless velar fricative \\ 
y & \unif{0079} & \textsc{latin small letter y} & close front rounded \\ 
ʏ & \unif{028F} & \textsc{latin letter small capital y} & lax close front rounded \\ 
z & \unif{007A} & \textsc{latin small letter z} & voiced alveolar fricative \\ 
ʐ & \unif{0290} & \textsc{latin small letter z with retroflex hook} & voiced retroflex fricative \\ 
ʑ & \unif{0291} & \textsc{latin small letter z with curl} & voiced alveolo-palatal fricative \\ 
ʒ & \unif{0292} & \textsc{latin small letter ezh} & voiced postalveolar fricative \\ 
ʔ & \unif{0294} & \textsc{latin letter glottal stop} & voiceless glottal plosive \\ 
ʕ & \unif{0295} & \textsc{latin letter pharyngeal voiced fricative} & voiced pharyngeal fricative \\ 
ʡ & \unif{02A1} & \textsc{latin letter glottal stop with stroke} & epiglottal plosive \\ 
ʢ & \unif{02A2} & \textsc{latin letter reversed glottal stop with stroke} & voiced epiglottal fricative \\ 
ǀ & \unif{01C0} & \textsc{latin letter dental click} & voiceless dental click \\ 
ǁ & \unif{01C1} & \textsc{latin letter lateral click} & voiceless alveolar lateral click \\ 
ǂ & \unif{01C2} & \textsc{latin letter alveolar click} & voiceless palatoalveolar click \\ 
ǃ & \unif{01C3} & \textsc{latin letter retroflex click} & voiceless (post)alveolar click \\ 
ʘ & \unif{0298} & \textsc{latin letter bilabial click} & voiceless bilabial click \\ 
β & \unif{03B2} & \textsc{greek small letter beta} & voiced bilabial fricative \\ 
θ & \unif{03B8} & \textsc{greek small letter theta} & voiceless dental fricative \\ 
χ & \unif{03C7} & \textsc{greek small letter chi} & voiceless uvular fricative \\
\end{xtabular}
\end{center}