### 2015-06-23 tables/unicode_valid.tex

\tablecaption{Additional characters for valid-IPA with Unicode encodings}\label{tab:ipa_lax}
 \tablefirsthead{
   \toprule
   & Code & Unicode name & Phonetic description \\ 
   \midrule
   }
 \tablehead{
   \multicolumn{4}{c}{
     \captionsize\tablename\ \thetable{} 
     Additional characters for valid-IPA with Unicode encodings --- \textit{continued}
     } \\
   \toprule
   & Code & Unicode name & Phonetic description \\ 
   \midrule
   }
 \tabletail{
   \bottomrule
   \multicolumn{4}{r}{
     \captionsize\textit{continued on next page}
     } \\
   }
 \tablelasttail{\bottomrule}

\begin{center}
\begin{xtabular}{ l L{1.2cm} L{6.1cm} L{3.5cm} }
\dia{030A} & \unif{030A} & \textsc{combining ring above} & voiceless (above) \\
g & \unif{0067} & \textsc{latin small letter g} & voiced velar plosive \\
\dia{030B} & \unif{030B} & \textsc{combining double acute accent} & extra high tone \\
\dia{0301} & \unif{0301} & \textsc{combining acute accent} & high tone \\
\dia{0304} & \unif{0304} & \textsc{combining macron} & mid tone \\
\dia{0300} & \unif{0300} & \textsc{combining grave accent} & low tone \\
\dia{030F} & \unif{030F} & \textsc{combining double grave accent} & extra low tone \\
\dia{0302} & \unif{0302} & \textsc{combining circumflex accent} & falling \\
\dia{030C} & \unif{030C} & \textsc{combining caron} & rising \\
\dia{1DC4} & \unif{1DC4} & \textsc{combining macron-acute} & high rising \\
\dia{1DC5} & \unif{1DC5} & \textsc{combining grave-macron} & low rising \\
\dia{1DC6} & \unif{1DC6} & \textsc{combining macron-grave} & low falling \\
\dia{1DC7} & \unif{1DC7} & \textsc{combining acute-macron} & high falling \\
\dia{1DC8} & \unif{1DC8} & \textsc{combining grave-acute-grave} & rising-falling \\
\dia{1DC9} & \unif{1DC9} & \textsc{combining acute-grave-acute} & falling-rising \\ 
\dia{035C}{\large\fontspec{CharisSIL}◌} & \unif{035C} & \textsc{combining double breve below} & tie bar (below) \\
\end{xtabular}
\end{center}