### 2015-09-29 book/localmetadata.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                                              %%%
%%%                 Metadata                     %%%
%%%          fill in as appropriate              %%%
%%%                                              %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{The Unicode Cookbook \linebreak for Linguists}
\subtitle{Managing writing systems using orthography profiles}
\BackTitle{Change backtitle in  localmetadata.tex}
\BackBody{Change backbody in  localmetadata.tex}
\dedication{Change dedication in localmetadata.tex}
\typesetter{Change typesetter in localmetadata.tex}
\proofreader{Change proofreaders in localmetadata.tex}
\author{Steven Moran \& Michael Cysouw}
\renewcommand{\lsISBN}{978--3--000000--00--0}                     
\renewcommand{\lsSeries}{dummyseries} 
% use lowercase acronym, e.g. sidl, eotms, tgdi
\renewcommand{\lsSeriesNumber}{99} 
%will be assigned when the book enters the proofreading stage
\renewcommand{\lsURL}{http://langsci-press.org/catalog/book/0} 
% contact the coordinator for the right number