### 2015-09-29 book/tables/unicode_additions.tex

\tablecaption{Additions to widened-IPA with Unicode encodings}\label{tab:ipa_additions}
 \tablefirsthead{
   \toprule
   & Code & Unicode name & Phonetic description \\ 
   \midrule
   }
 \tablehead{
   \multicolumn{4}{c}{
     \captionsize\tablename\ \thetable{} 
     Additions to widened-IPA with Unicode encodings --- \textit{continued}
     } \\
   \toprule
   & Code & Unicode name & Phonetic description \\ 
   \midrule
   }
 \tabletail{
   \bottomrule
   \multicolumn{4}{r}{
     \captionsize\textit{continued on next page}
     } \\
   }
 \tablelasttail{\bottomrule}

\begin{center}
\begin{xtabular}{ l L{1.1cm} L{6.1cm} L{3.5cm} }
‼ & \unif{203C} & \textsc{double exclamation mark} & retroflex click \\
\charis{ᶑ} & \unif{1D91} & \textsc{latin small letter d with hook and tail} & voiced retroflex implosive \\
\dia{0348} & \unif{0348} & \textsc{combining double vertical line below} & fortis \\
\dia{0349} & \unif{0349} & \textsc{combining left angle below} & lenis \\
\dia{0353} & \unif{0353} & \textsc{combining x below} & frictionalized \\
\dia{032E} & \unif{032E} & \textsc{combining breve below} & derhoticized \\
\dia{0347} & \unif{0347} & \textsc{combining equals sign below} & non-sibilant \\
\dia{02C0} & \unif{02C0} & \textsc{modifier letter glottal stop} & glottalized \\
\diareverse{02B1} & \unif{02B1} & \textsc{modifier letter small h with hook} & voiced pre-aspirated \\
\dia{1D31} & \unif{1D31} & \textsc{modifier letter capital e} & epilaryngeal phonation \\
\end{xtabular}
\end{center}