### 2015-10-17 book/chapters/codetesting.tex
testing code inclusion

\chapter{Testing code inclusing}

Trying to include code through knitr. Different languages can be easily 
included!

<<bash-test, engine='bash'>>=
echo 'this is a simple bash example'
ls | wc
@

<<>>=
'R works out of the box'
(x <- 3+4)
Sys.getlocale("LC_CTYPE")
'aɽɮz'
@

You can refer to variables by using \textbackslash Sexpr{}. 
I think \Sexpr{x} is an interesting number!

<<python-test, engine='python'>>=
'python also works!'
print('bla' + 'bla')
print(3+4)
@

You have to add the option `engine.paht='python3'' to get Python r

<<python3-test, engine='python', engine.path='python3'>>=
print('python3 needs an extra tag')
uni = 'aɽɮz'
print(uni)
@

Testing: does referencing to variables also work in Python? It does give an error, so no... !