### 2015-10-18 book/chapters/codetesting.tex
some more knitr testing

\chapter{Testing code inclusion}

Trying to include code through knitr. 
Different languages can be easily 
included!

There is a bug for non-R engines: the fontsize settings and line spaceing are not correct. I have 
filed an issue, hopefully this will be fixed soon.

For syntax colouring, install the package `highlight' and for nice tables 
install the package `xtable'.

<<bash-test, engine='bash'>>=
echo 'this is a simple bash example'
echo 'this is a simple bash example'
ls | wc
@

<<>>=
'this is a simple R example'
test = 3+4
test
@

<<>>=
library(qlcData)
example  = 'AΑА'
profile = write.profile(example)
@

<<Rtest2, tidy = TRUE, results = 'asis', echo = FALSE>>=
library(xtable)
print(xtable(profile, caption = 'test', label = 'bla')
  , include.rownames = FALSE
  , booktabs = TRUE
  , size = 'small'
  , table.placement = 'H'
  )
@

You can refer to variables by using \textbackslash Sexpr{}. 
For example, the length of the example string is \Sexpr{nchar(example)}.
Crossreferencing also works, see Table~\ref{bla}.

<<python-test, engine='python'>>=
print('bla' + 'bla')
print(3+4)
@

You have to add the option ``engine.path='python3'' to get Python 3.

<<python3-test, engine='python', engine.path='python3'>>=
print('python3 needs an extra tag')
uni = 'aɽɮz'
print(uni)
@

Testing: does referencing to variables also work in Python? It does give an error, so no... !