### 2016-06-22 book/chapters/profiles_examples.tex
minor changes

\section{Available implementations}
\label{implementations}

To illustrate the practical application of orthography profiles, we have
implemented two different versions of the above specifications, one in Python
and one in R. These two implementations have rather different implementation
histories, and we are not 100\% sure that they will in all situations give the
same results. Also the perfomance with larger datasets differs, and the code is
not always as clean as we would like it to be. In sum, the two implementations
should be considered as `proof of concept' and not as the final word on the
practical application of the specifications above. In our own experience, the
current implementations are sufficiently fast and stable to be useful for
academic practice (e.g. checking data consistency, or analyzing and
transliterating small to medium sized data sets), but they should probably
better not be used for full-scale industry applications.

\subsection*{Installing the R implementation}

The R implementation is available in the package \texttt{qlcData}, which is 
directly available from the central R repository CRAN (Comprehensive R Archive 
Network). The R software environment itself has to be downloaded from its 
website.\footnote{\url{https://www.r-project.org}} After starting the included 
R program, the \texttt{qlcData} package for dealing with orthography profiles can be 
simply installed as follows:

<<eval=FALSE>>==
# download and install the qlcData software
install.packages('qlcData') 
# load the software, so it can be used
library(qlcData) 
@

\noindent The version available through CRAN is the most recent stable version.
To obtain the most recent bug-fixes and experimental additions, please use the
development versions, which is available on
GitHub.\footnote{\url{http://github.com/cysouw/qlcData}} This development
version can be easily installed using the github-install helper software from the
\texttt{devtools} package.

<<eval=FALSE>>==
# download and install helper software
install.packages('devtools') 
# install the qlcData package from GitHub
devtools::install_github('cysouw/qlcData')
# load the software, so it can be used 
library(qlcData) 
@

\noindent Inside the \texttt{qlcData} package, there are two functions for
orthography processing, \texttt{write.profile} and \texttt{tokenize}. R includes
help files with illustrative examples, and also a so-called `vignette' with
explanations and examples.

<<eval=FALSE>>==
# view help files
help(write.profile)
help(tokenize)
# view vignette with explanation and examples
vignette('orthography_processing')
@

\noindent Basically, the idea is to use \texttt{write.profile} to produce a
basic orthography profile from some data and then \texttt{tokenize} to apply the
(possibly edited) profile on some data, as exemplified in the next section. This
can of course be performed inside R, but additionally there are two more
interfaces to the R code supplied in the \texttt{qlcData} package: (i) bash
executables and (ii) `shiny' webapps.

The bash executables are little files providing an interface to the R code that
can be used in a shell on a UNIX-alike machine. The location of these
executables is a bit hidden away by the install procedure of R packages. The
location can be found by the following command in R. These executables can be 
used from here, or they can be linked and/or copied to any location as wanted.

<<eval=FALSE>>==
# show the path to the bash executables
file.path(find.package('qlcData'), 'exec')
@

\noindent For example, a good way to use the executables in a terminal is to
make softlinks (using \texttt{ln}) from the executables to a directory in your
PATH, e.g. to \texttt{/usr/local/bin/}. The two executables are named
\texttt{tokenize} and \texttt{writeprofile}

<<link, engine='bash', size='scriptsize', tidy=FALSE, eval=FALSE>>==
# make two softlinks to the R executables in bash
ln -is `Rscript -e 'cat(file.path(find.package("qlcData"),
 "exec", "tokenize"))'` /usr/local/bin
ln -is `Rscript -e 'cat(file.path(find.package("qlcData"),
 "exec", "writeprofile"))'` /usr/local/bin
@

\noindent After this addition it should be possible to access the \texttt{tokenize}
function from the shell. Try \texttt{tokenize --help} to test the functionality.

<<help, size='tiny', engine='bash'>>=
tokenize --help
@

To make the functionality even more accessible, we have prepared webapps with 
the \texttt{shiny} framework for the R functions. These webapps are available 
online at \url{TODO}. The webapps are also included inside the \texttt{qlcData} 
package and can be started with the following helper function:

<<eval=FALSE>>==
launch_shiny('tokenize')
@

\subsection*{Installung the Python implementation}

TO DO

\section{Abstract examples}

\subsection*{Error reporting}

In the following example, it looks as if we have two identical strings,
\texttt{AABB}. However, this is just an surface-impression delivered by the
current font, which renders Latin and Cyrillic capitals identically. We can
identify this problem when we produce an orthography profile from the strings.
Using here the R implementation of orthography profiles, we first assign the two
strings to a variable \texttt{test}, and then produce an orthography profile
with the function \texttt{write.profile}. As it turns out, some of the letters 
are Cyrillic.

<<>>=
test <- c('AABB', '\u0041\u0410\u0042\u0412')
write.profile(test)
@

\noindent The working of error-message reporting can also nicely be illustrated with this
example. Suppose we made an orthography profile with just the Latin <A> and <B> 
as possible graphemes, then this profile would not be sufficient to tokenize the 
strings. There are graphemes in the data that are not in the profile, so the 
tokenization produces an error, which can be used to fix the encoding. In the 
example below, we can see that the Cyrillic encoding is found in the second 
string of the \texttt{test} input.

<<>>=
test <- c('AABB', '\u0041\u0410\u0042\u0412')
tokenize( test, profile = c('A', 'B') )
@

\subsection*{Different ways to write a profile}

The function \texttt{write.profile} can be used to prepare a skeleton for an
orthography profile from some data. The preparation of an orthography profile
from some data might sound like a trivial problem, but actually there are
various different ways in which strings can be separated into graphemes by
\texttt{write.profile}. Consider the following example string. The default
settings of \texttt{write.profile} separates the string into Unicode graphemes
according to grapheme clusters (`user-perceived characters', see
Section~\ref{the-unicode-approach}).

<<>>=
example <- '\u00d9\u00da\u00db\u0055\u0300\u0055\u0301\u0055\u0302'
profile_1 <- write.profile(example)
@

<<results = 'asis', echo = FALSE>>=
colnames(profile_1) <- c('Gr.', 'Freq.', 'Codepoint', 'Unicode Name')
print(xtable(profile_1
        , caption = 'Profile 1 (default settings, splitting grapheme clusters)'
        , label = 'tab:profile1'
        )
  , include.rownames = FALSE
  , booktabs = TRUE
  , size = 'scriptsize'
  , table.placement = 'H'
  )
@

\noindent Some of these graphemes are single codepoints, others are combinations
of two codepoints. By specifying the separator as empty, it is possible to split
the string into codepoints.

<<>>=
profile_2 <- write.profile(example, sep = "")
@

<<results = 'asis', echo = FALSE>>=
colnames(profile_2) <- c('Grapheme', 'Frequency', 'Codepoint', 'Unicode Name')
print(xtable(profile_2
        , caption = 'Profile 2 (splitting by codepoints)'
        , label = 'tab:profile2'
        )
  , include.rownames = FALSE
  , booktabs = TRUE
  , size = 'scriptsize'
  , table.placement = 'H'
  )
@

\noindent Some characters look identical, although they are encoded differently. 
Unicode offers different ways of normalization, which can be invoked here as 
well (see Section~\ref{pitfall-canonical-equivalence}).

<<>>=
# after NFC normalization unicode codepoints have changed
profile_3 <- write.profile(example, normalize = "NFC", sep = "")
# NFD normalization gives yet another structure of the codepoints
profile_4 <- write.profile(example, normalize = "NFD", sep = "")
@

<<results = 'asis', echo = FALSE>>=
colnames(profile_3) <- c('Grapheme', 'Frequency', 'Codepoint', 'Unicode Name')
print(xtable(profile_3
        , caption = 'Profile 3 (splitting by NFC codepoints)'
        , label = 'tab:profile3'
        )
  , include.rownames = FALSE
  , booktabs = TRUE
  , size = 'scriptsize'
  , table.placement = 'H'
  )
@

<<results = 'asis', echo = FALSE>>=
colnames(profile_4) <- c('Grapheme', 'Frequency', 'Codepoint', 'Unicode Name')
print(xtable(profile_4
        , caption = 'Profile 4 (splitting by NFD codepoints)'
        , label = 'tab:profile4'
        )
  , include.rownames = FALSE
  , booktabs = TRUE
  , size = 'scriptsize'
  , table.placement = 'H'
  )
@

\noindent It is important to realize that for Unicode grapheme definitions, NFC 
and NFD normalization are equivalent. This can be shown by normalizing the 
example in either NFC or NFD, but using default separation in 
\texttt{write.profile}. To be precise, default separation means setting \texttt{sep = NULL} 
but that has not be added explicitly.

<<>>=
# note that NFC and NFD normalization are identical
# for unicode grapheme definitions
profile_5 <- write.profile(example, normalize = "NFD")
profile_6 <- write.profile(example, normalize = "NFC")
@

<<results = 'asis', echo = FALSE>>=
colnames(profile_5) <- c('Gr.', 'Freq.', 'Codepoint', 'Unicode Name')
print(xtable(profile_5
        , caption = 'Profile 5 (splitting by graphemes after NFD)'
        , label = 'tab:profile5'
        )
  , include.rownames = FALSE
  , booktabs = TRUE
  , size = 'scriptsize'
  , table.placement = 'H'
  )
@

<<results = 'asis', echo = FALSE>>=
colnames(profile_6) <- c('Gr.', 'Freq.', 'Codepoint', 'Unicode Name')
print(xtable(profile_6
        , caption = 'Profile 6 (splitting by graphemes after NFC)'
        , label = 'tab:profile6'
        )
  , include.rownames = FALSE
  , booktabs = TRUE
  , size = 'scriptsize'
  , table.placement = 'H'
  )
@

\noindent Note that these different profiles can also be produced using the bash 
executable \texttt{writeprofile} (see \ref{implementations} for notes on how to 
install the bash executable). Exactly this example is also included in the help 
file of the executable.

<<writehelp, size='tiny', engine='bash'>>=
writeprofile --help
@

\subsection*{Using an orthography profile skeleton}