### 2017-01-20 book/chapters/preface.tex
cleanup

\chapter{Preface}
\label{preface}

% ==========================
% \section{Preface}
% \label{preface}
% ==========================

% Very early attmept at notion for linguistic prosody: Joshua Steele (Prosodia Rationalis, 1779)

% Early phoneticians: Richard Lepsius (standard alphabet 1854, 1855), Isaac Pitman (Stenographic Soundhand 1837, Phonotypy 1844, English Phonotypic Alphabet 1845 with Ellis), Alexander J. Ellis (English Phonotypic Alphabet 1845), and Alexander Melville Bell (visibile speech 1864, 1867)

% Later phoneticians: Henry Sweet (Romic alphabet, aka Romic reform, in A Handbook of Phonetics 1877 - later becomes basis for IPA), L'abbé Jean-Pierre Rousselot ("founder of experimental phonetics"), Eduard Sievers, Paul Passy (founded the Phonetic Teachers' Association in 1886, later became IPAssociation), and Jens Otto Harry Jesperson

% Instrumental measurements: Ludimar Hermann (used the phonograph to test theories of vowel production; coined formant)

In the 1840s, several researchers and teachers began developing notation to
accurately transcribe articulatory gestures and their acoustic correlates. Their
goal was not only transcription, they wanted to use their notation for spelling
reform (especially English), for teaching literacy and foreign languages, and in
particular for developing writing systems for unwritten languages.

At first, the challenge of devising a universal phonetic notation seemed
straightforward: normalize the relationship between graphemes and sounds (<c> in
`cake' -> /k/; <c> in `cell' -> /s/) to remove inconsistencies introduced by the
historical divergence of written and spoken forms (e.g.\ in English and French
orthographies), and keep adding new symbol-sound mappings until all sounds are
accounted for.

% These early transcription system pioneers found success with the first few languages that they encountered. Grapheme replacement and normalization of symbols across different orthographies seemed to work. But as they were confronted with more articulations and different notational devices, early phoneticians encountered a wealth of potentially important data that they had not expected.

This approach was successful for the first few languages investigated. But as
the early phoneticians were confronted with more and more articulations, they
encountered a wealth of data that they had not expected. The in-depth
exploration of articulatory phonetics led to the discovery that the number of
possible spoken sounds is infinite. \cite{Hockett1995} notes:

\begin{quote}

But as they observed articulation more and more closely, and as
they took more and more languages into consideration, the number of sounds they
could discriminate and the number of symbols needed to denote them both
increased seemingly without limit. In fact, as early as 1848, Ellis asserted
that in principle the number of different sounds is endless. With the
development of instrumental phonetics, largely by L'Abbé Rousselot, it soon
turned out that, given sufficiently accurate measuring devices, that is in fact
the case.

\end{quote}

% Hockett 1995: "And so the phoneticians began to develop notations that would show only the constant and distinctive features and let the transient and predictable ones be provided for by cover statements"

% The phonemic principle emerged and the model for it was alphabetic writing.

The first instrumental measurements in the 1870-80s showed unequivocally that
spoken language is comprised of sounds waves; it is dynamic and not discrete.
Given the fluid nature of the speech stream, sounds differ from one utterance to
the next and the same sound may differ in different phonetic environments.
However, many of these alterations are predictable in the context of their
environments and they need not necessarily form distinctive lexical contrasts.

These findings left the early phoneticians in a conundrum given the aims of
developing precise phonetic notion, but also practically-minded alphabetic-based
writing. It is in this vein that the phonetic and phonemic distinction was
introduced very early into transcription. The etic vs emic principle took hold,
in essence, even before the dichotomy was widely accepted and before terminology
for linguistic distinctiveness was introduced and adopted.

A transcription system that showed both contrastive and non-contrastive elements
emerged, taking form notably in the International Phonetic Alphabet (which
itself has been updated from time-to-time over the last nearly 120 years). The
association between phonetic and phonemic would be maintained in the
International Phonetic Alphabet and through the various instantiations of
phonological theory, which developed on the work by early phoneticians.

The model for phonemic contrast was alphabetic writing, in particular shallow
orthography, i.e.\ one-to-one mappings between sound and symbol. Traditional
alphabetic orthography, however, did not scale to precise phonetic notation, so
the former was sprinkled with diacritics to denote non-contrastive phenomena.
This dichotomy between phonemic and phonetic and its implementation in writing
and consistent encoding causes many issues -- linguistic, theoretic, and most
recently technological -- which we bring to light in this book.