### 2018-02-04 book/chapters/practical_recommendations.tex
addresses issue #17

\chapter{Practical recommendations}
\label{practical-recommendations}

% A section which is missing in something called a "Cookbook" would be
% $ practical recommendations on how to input Unicode characters. There are
% various character selection tools, shortcuts on the keyboard, the
% shapecatcher website references at several places, or the Wikipedia
% lists of glyphs and fileformat.info. Having all this in one section
% would be handy for the user. It is of course unrelated to the
% orthography profiles, but I imagine that many people will use this book
% as a primer on IPA+Unicode and actually disregard the last two chapters.
% For this group, such a summary would be useful.

In this chapter, we provide some practical recommendations for quickly finding information about the Unicode Standard, the International Phonetic Alphabet (IPA), and tools and resources for working with each. This chapter is meant to be a short guide for novice users who are not interested in the programmatic aspects presented in the next two chapters. Instead, we aim here at ordinary working linguists who want to read about the basics and who want to easily insert special characters into their digital documents and applications.

\section{Unicode}
We discussed the Unicode Consortium's approach to computationally encoding writing systems in detail in Section \ref{the-unicode-approach}. The common pitfalls that we have encountered when using the Unicode Standard are discussed in detail in Chapter \ref{unicode-pitfalls}. Together these chapters provide users with an in-depth background about the hurdles they may encounter when using the Unicode Standard for encoding linguistic data or for developing multilingual applications. The provided inforation is by no means exhaustive and it is aimed at both linguists and programmers. For general background information about Unicode and character encodings, try these resources:

\begin{itemize}
	\item \url{http://www.unicode.org/standard/WhatIsUnicode.html}
	\item \url{https://en.wikipedia.org/wiki/Unicode}
	\item \url{https://www.w3.org/International/articles/definitions-characters/}
\end{itemize}

% \section{Unicode character pickers}
For practical purposes, users need a way to insert special characters (i.e.\ characters that are not easily entered via their keyboards) into documents and software applications. There are a few basic approaches for inserting special characters. One way is to use software-specific functionality, when it is available. For example, Microsoft Word has an insert-special-symbol-or-character function that allows users to scroll through a table of special characters across different scripts. Special characters can be then inserted into the document by clicking on them. Another way is to install a system-wide application for special character insertion. We have long been fans of the PopChar application from Ergonis Software, which is a small program that can insert Unicode characters.\footnote{\url{http://www.ergonis.com/products/popcharx/}}

There are also web-based so-called Unicode character pickers available through the browser that allow for the creation and insert of special characters, which can then be copied \& pasted into documents or software applications. For example:

\begin{itemize}
	\item \url{https://unicode-table.com/en/}
	\item \url{https://r12a.github.io/pickers/}
\end{itemize}

Yet another option for special character insertion includes operating system-specific shortcuts. For example on the Mac, holding down a key on the keyboard for a second, say <u>, triggers a pop up with the options <û, ü, ù, ú, ū> which can then be chosen by number (1--5). This method is convenient for occasionally inserting special characters, but it is burdensome for inserting many characters.

Therefore, there are many applications for installing language-specific or script-specific keyboards. These programs typically override keys or keystrokes on the user's keyboard and therefore allow them to quickly keyboard the insertion of specal characters (once the layout of the new keyboard is mastered). Many operating systems provide such keyboards as part of the base operating system, e.g.\ Microsoft Windows, Linux, and OS X, have special character Unicode eyboards pre-installed. Seach for how to access these pre-installed options on your operating system.

Lastly, there are third party applications that provide virtual keyboards. They can be language-specific or devoted specifically to IPA. Two popular programs are:

\begin{itemize}
	\item \url{https://keyman.com/}
	\item \url{http://scripts.sil.org/ipa-sil_keyboard}
\end{itemize}

\section{IPA}
In Chapter \ref{the-international-phonetic-alphabet} we described in detail the history of the International Phonetic Alphabet and how it became encoded in the Unicode Standard. In Chapter \ref{ipa-meets-unicode} we describe the resulting pitfalls from their marriage. These two chapters provide a very detailed overview of the challenges that users face when working with the two standards.

For general information about the International Phonetic Alphabet, the standard text is the \textit{Handbook of the International Phonetic Association: A Guide to the Use of the International Phonetic Alphabet} \cite{IPA2007}. It described in detail the principles and premises of the IPA that we summarize in Section \ref{IPApremises-principles}. The Association makes available information about itself  online\footnote{\url{https://www.internationalphoneticassociation.org/}} and it also makes available the International Phonetic Alphabet in chart form.\footnote{\url{https://www.internationalphoneticassociation.org/content/ipa-chart}} Wikipedia also has a comprehensive article about the IPA.\footnote{\url{https://en.wikipedia.org/wiki/International_Phonetic_Alphabet}}

There are several good Unicode IPA character pickers available on the Web, including:

\begin{itemize}
	\item \url{https://r12a.github.io/pickers/ipa/}
	\item \url{https://westonruter.github.io/ipa-chart/keyboard/}
	\item \url{http://ipa.typeit.org/}
\end{itemize}

\noindent Various linguistics departments also provide information about IPA fonts, software, and inserting Unicode IPA characters. Some good ones are:

\begin{itemize}
	\item \url{http://www.phon.ucl.ac.uk/resource/phonetics/}
	\item \url{https://www.york.ac.uk/language/current/resources/freeware/ipa-fonts-and-software/}
\end{itemize}

Regarding fonts that display Unicode IPA correctly, many linguists turn to the IPA Unicode fonts developed by SIL Internnational. The complete SIL font list is available online.\footnote{\url{http://scripts.sil.org/SILFontList}} There is also a page that describes IPA transcription using the SIL fonts and provides an informative discussion on deciding which font to use.\footnote{\url{http://scripts.sil.org/ipahome}} Traditionally, IPA fonts popular with linguists were created and maintained by SIL International, so it is often the case in our experience that we encounter linguistics data in legacy IPA fonts, i.e.\ pre-Unicode fonts such as SIL IPA93.\footnote{\url{http://scripts.sil.org/FontFAQ_IPA93}} SIL International does a good job of describing how to convert from legacy IPA fonts to Unicode IPA. By far the most popular fonts by SIL International are Doulos SIL and Charis SIL:

\begin{itemize}
	\item \url{https://software.sil.org/doulos/}}
	\item \url{https://software.sil.org/charis/}
\end{itemize}

\section{Advanced topics}
If you have made it this far, and you are eager to know more about the technological aspects of the Unicode Standard and how they relate to software programming, we recommend two light-hearted blog posts on the topic. The classic blog post about what programmers should know about the Unicode Standard is Joel Spolsky's ``The Absolute Minimum Every Software Developer Absolutely, Positively Must Know About Unicode and Character Sets (No Excuses!)''.\footnote{\url{https://www.joelonsoftware.com/2003/10/08/the-absolute-minimum-every-software-developer-absolutely-positively-must-know-about-unicode-and-character-sets-no-excuses/}} A more recent blogpost, with a bit more of the technical details, is by David C. Zentgraf and is titled, ``What Every Programmer Absolutely, Positively Needs To Know About Encodings And Character Sets To Work With Text''.\footnote{\url{http://kunststube.net/encoding/}} This post is aimed at software developers and uses the PHP language for examples.

Here are some online resources that we find particularly useful for finding more information about individual Unicode characters and also for converting between encodings:

\begin{itemize}
	\item \url{http://www.fileformat.info/}
	\item \url{https://unicodelookup.com/}
	\item \url{https://r12a.github.io/scripts/featurelist/}
	\item \url{https://r12a.github.io/app-conversion/}
\end{itemize}

For users of Python, see the standard documentation on how to use Unicode in your programming applications.\footnote{\url{https://docs.python.org/3/howto/unicode.html}} For \latex users, the TIPA package is useful for inserting IPA characters into your typeset documents. See these resources:

\begin{itemize}
	\item \url{http://www.tug.org/tugboat/tb17-2/tb51rei.pdf}
	\item \url{https://ctan.org/pkg/tipa}
	\item \url{http://ptmartins.info/tex/tipacheatsheet.pdf}
\end{itemize}

Lastly, we leave you with some Unicode humor for making it this far:

\begin{itemize}
	\item \url{https://xkcd.com/380/}
	\item \url{https://xkcd.com/1137/}
	\item \url{http://www.commitstrip.com/en/2014/06/17/unicode-7-et-ses-nouveaux-emoji/}
	\item \url{http://www.i18nguy.com/humor/unicode-haiku.html}
\end{itemize}