### 2018-06-11 book/langsci/seriesinfo/sidl-info.tex

%% -*- coding:utf-8 -*-
{\large Studies in Diversity Linguistics}

\bigskip

Chief Editor: Martin Haspelmath \\
Consulting Editors: Fernando Zúñiga, Peter Arkadiev, Ruth Singer, Pilar Valen­zuela

\bigskip

In this series:

\begin{enumerate}
\item Handschuh, Corinna. A typology of marked-S languages.
\item Rießler, Michael. Adjective attribution.
\item Klamer, Marian (ed.). The Alor-Pantar languages: History and typology.
\item Berghäll, Liisa. A grammar of Mauwake (Papua New Guinea).
\item Wilbur, Joshua. A grammar of Pite Saami.
\item Dahl, Östen. Grammaticalization in the North: Noun phrase morphosyntax in Scandinavian vernaculars.
\item Schackow, Diana.    A grammar of Yakkha.
\item Liljegren, Henrik. A grammar of Palula.
\item Shimelman, Aviva. A grammar of Yauyos Quechua.
\item Rudin, Catherine \& Bryan James Gordon (eds.). Advances in the study of Siouan languages and linguistics.
\item Kluge, Angela. A grammar of Papuan Malay.
\item Kieviet, Paulus. A grammar of Rapa Nui.
\item Michaud, Alexis. Tone in Yongning Na: Lexical tones and morphotonology.
\item Enfield, N.\,J (ed.).  Dependencies in language: On the causal ontology of linguistic systems .
\item Gutman, Ariel. Attributive constructions in North-Eastern Neo-Aramaic.
\item Bisang, Walter \& Andrej Malchukov (eds.). Unity and diversity in grammaticalization scenarios.
\item Stenzel, Kristine \& Patrizia Paggio (eds.). On this and other worlds: Voices from Amazonia.
\item Paggio, Patrizia and Albert Gatt (eds.). The languages of Malta.
\item Ilja A. Seržant \& Alena Witzlack-Makarevich (eds.). The diachronic typology of differential argument marking.

% \item Yakpo 
% \item Aguilar
% \item Riesberg
% \item di Garbo
% \item Floyd
% \item Haspelmath
% \item Veselinova
\end{enumerate}