### 2018-06-11 book/localmetadata.tex

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                                              %%%
%%%                 Metadata                     %%%
%%%          fill in as appropriate              %%%
%%%                                              %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{The Unicode Cookbook \linebreak for Linguists}
\subtitle{Managing writing systems using orthography profiles}
\BackTitle{Change backtitle in  localmetadata.tex}
\BackBody{Change backbody in  localmetadata.tex}
% \dedication{Change dedication in localmetadata.tex}
\typesetter{Change typesetter in localmetadata.tex}
\proofreader{Change proofreaders in localmetadata.tex}
\author{Steven Moran \and Michael Cysouw}
% \renewcommand{\lsISBNdigital}{978--3--000000--00--0}                     
% \renewcommand{\lsISBNhardcover}{978--3--000000--00--0}                     
\renewcommand{\lsSeries}{tmnlp}  
\renewcommand{\lsSeriesNumber}{10}  
\renewcommand{\lsID}{176}