### 2018-06-11 book/localmetadata.tex
add some of the proofreaders

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%                                              %%%
%%%                 Metadata                     %%%
%%%          fill in as appropriate              %%%
%%%                                              %%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{The Unicode Cookbook \linebreak for Linguists}
\subtitle{Managing writing systems using orthography profiles}
\BackTitle{Change backtitle in  localmetadata.tex}
\BackBody{Change backbody in  localmetadata.tex}
% \dedication{Change dedication in localmetadata.tex}
\typesetter{Change typesetter in localmetadata.tex}
\proofreader{Sandra Auderset, Aleksandrs Berdičevskis, Rosey Billington, Simon Cozens, Aniefon Daniel, Bev Erasmus, Amir Ghorbanpour, David Lukeš, Hugh Paterson III, Stephen Pepper, Katja Politt, Felix Rau, James Tauber, Luigi Talamo, Jeroen van de Weijer, Viola Wiegand, Alena Witzlack-Makarevich, and Esther Yap}
\author{Steven Moran \and Michael Cysouw}
\renewcommand{\lsISBNdigital}{978-3-96110-090-3 }                     
\renewcommand{\lsISBNhardcover}{978-3-96110-091-0}                     
\renewcommand{\lsSeries}{tmnlp}  
\renewcommand{\lsSeriesNumber}{10}  
\renewcommand{\lsID}{176}